---
title: "Download and wrangle hydro datasets"
format: 
  html:
    toc: true
    number-sections: true
    embed-resources: true
authors: Michelle Stern and Nick Rowlands
---

# Download data

```{r}
# Required packages to include
# install.packages("librarian")
librarian::shelf(leaflet, sf)
```

![[SCHISM boundary conditions](https://cadwrdeltamodeling.github.io/BayDeltaSCHISM/topics/baydelta_boundaries.html)](https://cadwrdeltamodeling.github.io/BayDeltaSCHISM/_images/open_boundaries.png){width="70%"}

```{r}
# SCHISM data

```

# Download other data

- Get instantaneous USGS data (for example, discharge sensor data). Start here: ?read_waterdata_continuous. If you only need the latest value, you can use the function: ?read_waterdata_latest_continuous.

- Get daily USGS data (for example, mean daily discharge). Start here: ?read_waterdata_daily. If you only need the latest value, you can use the function: ?read_waterdata_latest_daily.

- Get discrete USGS groundwater data. Start here: ?read_waterdata_field_measurements

- Get discrete water quality data from a cooperative service that integrates publicly available water-quality data from the USGS, EPA, and over 400 state, federal, tribal, and local agencies. Start here: ?readWQPdata

- Get USGS discrete water quality data. Start here: ?read_waterdata_samples

- Get metadata about USGS time series data, including instantaneous and daily data. Start here: ?read_waterdata_ts_meta

- Find Hydro Network-Linked Data Index (NLDI) data. Start here: ?findNLDI

- Get daily data statistics. Start here: ?read_waterdata_stats_por

- Add a USGS WaterData API token to your R environment. See: https://doi-usgs.github.io/dataRetrieval/articles/read_waterdata_functions.html#api-tokens

- Get the latest status on NWIS and other data access updates. See: https://doi-usgs.github.io/dataRetrieval/articles/Status.html https://api.waterdata.usgs.gov/ogcapi/v0/collections/monitoring-locations

```{r}
# https://github.com/DOI-USGS/dataRetrieval

```

# Visualize data

```{r}
# gather shapefiles for interactive map of Delta
delta_shapefile_url <- "https://utility.arcgis.com/usrsvcs/servers/3efc635b27344a3da989ca1e7108f5e0/rest/services/Boundaries/i03_LegalDeltaBoundary/MapServer/0/query?outFields=*&where=1%3D1&f=geojson"

# Read the GeoJSON directly from the URL
# geojsonio::geojson_read returns a Spatial object
delta_legal <- st_read(delta_shapefile_url, quiet = TRUE)
# Define projection
delta_legal <- st_transform(delta_legal, crs = 4326)
```

```{r}
# interactive map of Delta with leaflet
leaflet(delta_legal) %>%
  addTiles() %>%  # Add default OpenStreetMap tiles
  setView(lng = mean(st_coordinates(st_centroid(st_union(delta_legal)))[,1]),
          lat = mean(st_coordinates(st_centroid(st_union(delta_legal)))[,2]),
          zoom = 10) %>%
  addPolygons(
    color = "blue",
    weight = 1,
    opacity = 0.6,
    fillOpacity = 0.2,
    popup = ~as.character("Legal Delta") # can change to a shapefile column name
  )
```

# Merge data together

# Export to format needed for modeling
