---
title: "Grabbing Chlorophyll data"
format: html: embed-resources: TRUE
author: Emily Richardson, Shaela Noble, Spencer Fern (others tbd)

---
# This document shows steps that the 2026 NCEAS CHAB team is using to harmonize chlorophyll data

## Emily's Workspace

### ScienceBase
#### Will either need custom API or individual downloads


### Water data for the nation
#### will use https://github.com/DOI-USGS/dataRetrieval
```{r}
#load packages
library(dataRetrieval)
library(dplyr)
library(purrr)
library(sf)
library(stringr)
```
```{r}
# Identify Delta surface water sites
out_sf <- 
  dataRetrieval::read_waterdata_monitoring_location(
    state_name = "California", 
    hydrologic_unit_code = "18020163",
    agency_name = "U.S. Geological Survey",
    properties = c("monitoring_location_id", "site_type", "hydrologic_unit_code", "county_name", "agency_name"))

sites <- out_sf %>%
  mutate(lon = st_coordinates(.)[,1],
         lat = st_coordinates(.)[,2])

delta_sites <- sites %>% filter((between(lat, 37.585545, 38.566034)) &  between(lon, -122.292649, -121.131962))
```

```{r}
# Find chlorophyll-related parameter codes
all_pcodes <- read_waterdata_parameter_codes()
chl_pcodes <- all_pcodes %>% filter(str_detect(parameter_name, "Chl"))
#chl_pcodes <- chl_pcodes %>% filter(parameter_group_code == "BIO")
```

```{r}
sitelist <- delta_sites$monitoring_location_id
deltasite_metadata <- read_waterdata_ts_meta(
    monitoring_location_id = sitelist,
    properties = c("monitoring_location_id", "parameter_code", "begin_utc", "end_utc"))
```

```{r}
#filter out Delta sites that don't have potential chlorophyll-related data
delta_chl_sites <- deltasite_metadata %>% 
    filter(parameter_code == "32316")
#lu_delta_pcodes <- unique(delta_chl_sites$parameter_code)
lu_delta_chl_sites <- unique(delta_chl_sites$monitoring_location)
```

```{r}
mindate <- min(delta_chl_sites$begin_utc, na.rm = TRUE)
mindate <- as.Date(mindate)
mindate <- as.character(mindate)

maxdate <- max(delta_chl_sites$end_utc, na.rm = TRUE)
maxdate <- as.Date(maxdate)
maxdate <- as.character(maxdate)
```

```{r}
#BROKEN - too many dates need to figure out how to loop- Now grab the data (this is for mean daily sensor chl only in ug/l)
delta_chl_data <- dataRetrieval::read_waterdata_continuous(monitoring_location_id = lu_delta_chl_sites,
                              parameter_code = lu_delta_pcodes,
                              time = c(mindate, maxdate)
                              )
```

```{r}
summary(delta_chl_data)
```

```{r}
#Grab only daytime data and average daily

library(lubridate)

delta_chl_data$hour <- as.numeric(format(delta_chl_data$time, "%H"))

delta_chl_data$TOD <- ifelse(delta_chl_data$hour %in% c(8, 9, 10, 11, 12, 13, 14, 15, 16), "keep", "remove")
delta_chl_data$date <- as.Date(delta_chl_data$time)

daytime_delta_chl_data <- delta_chl_data %>% filter(TOD == "keep")

avg_daytime_delta_chl_data <- daytime_delta_chl_data %>%
  group_by(date, monitoring_location_id)%>%
  mutate(avg_fchl = mean(value)) %>%
  mutate(fchl_corr = 1.29*avg_fchl+0.33)%>%
  select(date, monitoring_location_id, fchl_corr, avg_fchl)%>%
  distinct()
```

```{r}
summary(avg_daytime_delta_chl_data)
```



## Shaela's Workspace

### Discrete WQ Data Package

#### Install, load package and read in data


```{r}
#devtools::install_github("InteragencyEcologicalProgram/discretewq")

library(discretewq)
library(dplyr)
library(sf)
```
```{r}
discrete_raw <- wq(
  Sources = c(
    "20mm",
    "Baystudy",
    "DJFMP",
    "DOP",
    "EDSM",
    "EMP",
    "FMWT",
    "NCRO",
    "SDO",
    "SKT",
    "SLS",
    "STN",
    "Suisun",
    "USBR",
    "USGS_CAWSC",
    "USGS_SFBS",
    "YBFMP"
  )
)

str(discrete_raw)
```

#### Data exploration
```{r}
#filter to chlorophyll
discrete_chla <- discrete_raw %>% 
    filter(!is.na(Chlorophyll))

#see which surveys have chlorophyll data
discrete_chla %>% 
    distinct(Source)
```

```{r}
#filter out any NA in coordinates columns and convert to a spatial object
discrete_chla_sf <- discrete_chla %>% 
    filter(!is.na(Latitude)) %>% 
    st_as_sf(coords = c('Longitude', 'Latitude'),
             crs = 4326,
             remove = FALSE)

test <- discrete_chla_sf %>% 
    filter(Source == "DOP") %>% 
    distinct(Latitude, Longitude, .keep_all = TRUE)
```



## Spencer's Workspace

### e.g., name of data source

## Running spatial data packages
```{r}
### Data wrangling packages:
library(readr)
library(dplyr)
library(tidyr)
library(here)

### Simple Features vector spatial data:
library(sf)

### Data visualization packages:
library(ggplot2)
library(leaflet)
library(ggspatial) 
library(prettymapr) ### ggspatial depends on but does not import prettymapr for internal functions
```

```{r}
chla_special_sf <-read_csv(here("data/model_chla_covars.csv"))

chla_stations <- read_csv(here("data/stations.csv"))
```

```{r}
chla_lat_long <- chla_stations %>%
    select(station, latitude, longitude) %>%
    mutate(latitude = round(latitude, 3), 
        longitude = round(longitude, 3)    
        ) %>%
    distinct()
chla_joined <- chla_special_sf %>%
    full_join(chla_lat_long, by = "station")

write.csv(chla_joined, "data/01_tidy_model_chla_covars.csv")
```


```{r}
chla_4326_sf <- st_as_sf(chla_joined,
                        coords = c('longitude', 'latitude'),
                        crs = 4326)
```


```{r}
plot(chla_4326_sf %>%
    select(chlorophyll))
```


## Making a Map with the Chlorophyll Stations

```{r}
ggplot(chla_4326_sf) +
    geom_sf(aes(fill = chlorophyll)) +
    labs(fill = "chlorophyll") +
    scale_fill_continuous(low = "blue", high =  "green") +
    theme_bw()
```


```{r}
chla_3857_sf <- st_transform(chla_4326_sf, 
                            crs = 3857)
```


```{r}
plot(chla_3857_sf %>%
        select(chlorophyll))
 ggplot(chla_3857_sf) +
     ggspatial::annotation_map_tile(type = "osm", zoom = 10, progress = 'none') +
       geom_sf(aes(fill = chlorophyll)) +
     labs(fill = "chlorophyll") +
     scale_fill_continuous(low = "blue", high =  "green") +
     theme_bw()
```
