---
title: "Grabbing Chlorophyll data"
format: html: embed-resources: TRUE
author: Emily Richardson, Shaela Noble, Spencer Fern (others tbd)

---
# This document shows steps that the 2026 NCEAS CHAB team is using to harmonize chlorophyll data

## Emily's Workspace

```{r}
#load packages
library(dataRetrieval)
library(dplyr)
library(purrr)
library(sf)
library(stringr)
library(DT)
library(lubridate)
library(DBI)
library(EDIutils)
library(sbtools)
```

### ScienceBase
```{r}
# Pulling ScienceBase data using the R package sbtools
## Defining my search area
delta_items <- query_sb_spatial(
  long = c(-122.292649, -121.131962),
  lat  = c(37.585545, 38.566034),
  limit = 100000
)

## Extract metadata
titles <- sapply(delta_items, function(x) x$title)

## Keep chlorophyll-related records
idx <- grepl(
  "chlorophyll|chlorophyll-a|phytoplankton",
  titles,
  ignore.case = TRUE
)

delta_chl <- delta_items[idx]
```

```{r}
#Read through returned titles to verify relevancy
print(delta_chl) #30 relevant datasets
#1 N - Little Holland Tract - These are enumeration data, not chl concentration
#2 N - Phytoplankton tally sheet - These are enumeration data, not chl concentration
#3 Y - Assessing spatial variability - Contains Chlorophyll data
#4 Y - Characterization of water - Contains Chlorophyll data
#5 Y - Assessing spatial variability - Contains Chlorophyll data
#6 Y - Assessment of nutrients - Contains Chlorophyll data
#7 Y - Cyanotoxin, chlorophyll-a, and - Contains Chlorophyll data
#8 N - Phytoplankton Species Composition - These are enumeration data, not chl concentration
#9 N - Aquatic reflectance data - These data are not within our study area (bounding box encompasses US)
#10 Y - Phytoplankton data for - Contains chl-a data from Freeport station (and from other sites throughout US)
#11 Y - Assessing spatial variability - Contains Chlorophyll data
#12 N - Water chemistry, vertical - Does contain chl data but collected during nutrient addition experiments, not monitoring.
#13 Y - Assessing spatial variability - Contains Chlorophyll data
#14 Y - Assessing spatial variability - Contains Chlorophyll data
#15 Y - Assessing spatial variability - Contains Chlorophyll data
#16 Y - Assessing spatial variability - Contains Chlorophyll data
#17 N - Daily Estimates of - Just Site metadata
#18 N - Daily Estimates of - CM sensor data, already pulling this
#19 Y - A national harmonized - Contains Chlorophyll data (US-wide, including the Delta)
#20 Y - Cyanotoxin, chlorophyll-a, and - Contains chl-a AND Qpcr!! data from Freeport station (and from other sites throughout US) - are these data the same as number 10?
#21 N - Phytoplankton data for - These are enumeration data, not chl concentration
#22 N - Phytoplankton Identification and - These are enumeration data, not chl concentration
#23 Y - Water quality, nutrient, - Contains Chlorophyll data
#24 N - Phytoplankton Species Composition - These are enumeration data, not chl concentration
#25 Y - Cyanotoxin, chlorophyll-a, and - Contains chl-a and qpcr
#26 N - Daily Estimates of Chlorophyll - Just approval status
#27 N - Daily Estimates of - CM sensor data, already pulling this
#28 N - Phytoplankton data for - These are enumeration data, not chl concentration
#29 N - Phytoplankton and Cyanotoxin - These are National Parks data
#30 Y - Assessing spatial variability - Contains Chlorophyll data (but provisional!)
```

```{r}
#The data I want are associated with the following numbers: 3,4,5,6,7,10,11,13,14,15,16,19,20,23,25,30
##create list of item ids
ids <- c(
  "64fa6804d34ed30c205504a6",
  "5f6ce12a82ce38aaa24801be",
  "5f24747f82cef313ed971974",
  "5f5fe6eb82ce3550e3bff4a5",
  "601827aad34edf5c66efffc8",
  "601025fcd34e162231fecfca",
  "5f7dee6482ce1d74e7dd9da0",
  #"62183869d34ec739b2dd292f", Bay Based and too large
  "6596fe37d34e3265ab158d60",
  "5f284f1782cef313ed9df50e",
  "5fa096b1d34e198cb7919341",
 # "638f5472d34ed907bf7c8f23", #this one is too big to store in git. It is just freeport and we catch that with usgs CM, so skipping.
  "628cc9c2d34ef70cdba3bbff",
  "64ee7f82d34e4b6d8ab71438",
  "5c13d214e4b006c4f8532a17",
  "66427d79d34e1955f5a41cf0"
)

files_list <- lapply(ids, item_list_files)

download_dir <- "data/raw/sciencebase"
#dir.create(download_dir, recursive = TRUE, showWarnings = FALSE)

for (id in ids) {
  item_dir <- file.path(download_dir, id)

  dir.create(item_dir, showWarnings = FALSE)

  item_file_download(
    sb_id = id,
    dest_dir = item_dir
  )
}

rm(delta_chl, delta_items, files_list, id, ids, idx, item_dir, titles, download_dir)
```

```{r}
#I noticed some data that I had identified via keyword searching for "chlorophyll sacramento" are not in the above data pull. Script to add them:
ids <- c(
  "5966abe6e4b0d1f9f05cf551",
  "5e7ceb88e4b01d5092751c18",
  "5f1f44c582cef313ed8ec753",
  "5f89cbea82ce3241878937f6",
  "5d3f1ae2e4b01d82ce8d90e0",
 # "5f452e8782ce4c3d12251702", #this one is too big to store in git, large fies didn't contain chl anyways so removed.
  "64248ee5d34e370832fe343d",
  "65b18ae7d34e36a39044c493",
  "5f5ff87582ce3550e3bff547"
)

files_list <- lapply(ids, item_list_files)

download_dir <- "data/raw/sciencebase"

for (id in ids) {
  item_dir <- file.path(download_dir, id)

  dir.create(item_dir, showWarnings = FALSE)

  item_file_download(
    sb_id = id,
    dest_dir = item_dir
  )
}
rm(files_list, id, ids, idx, item_dir, titles, download_dir)
```

```{r}
#remote: error: File data/raw/sciencebase/638f5472d34ed907bf7c8f23/national_chlorophyll_corrected_chlorophyll_a_data.csv is 106.56 MB; this exceeds GitHub's file size limit of 100.00 MB
#remote: error: File data/raw/sciencebase/5f452e8782ce4c3d12251702/FTS_acs_ag.csv is 118.54 MB; this exceeds GitHub's file size limit of 100.00 MB
#remote: error: File data/raw/sciencebase/5f452e8782ce4c3d12251702/FTS_acs_apg.csv is 117.61 MB; this exceeds GitHub's file size limit of 100.00 MB
#remote: error: File data/raw/sciencebase/5f452e8782ce4c3d12251702/FTS_acs_cg.csv is 117.94 MB; this exceeds GitHub's file size limit of 100.00 MB

```

### Water data for the nation (USGS)
```{r}
# Identify Delta surface water sites
out_sf <- 
  dataRetrieval::read_waterdata_monitoring_location(
    state_name = "California", 
    hydrologic_unit_code = "18020163",
    agency_name = "U.S. Geological Survey",
    properties = c("monitoring_location_id", "site_type", "hydrologic_unit_code", "county_name", "agency_name"))

sites <- out_sf %>%
  mutate(lon = st_coordinates(.)[,1],
         lat = st_coordinates(.)[,2])

delta_sites <- sites %>% filter((between(lat, 37.585545, 38.566034)) &  between(lon, -122.292649, -121.131962))

rm(out_sf, sites)
```

```{r}
sitelist <- delta_sites$monitoring_location_id
deltasite_metadata <- read_waterdata_ts_meta(
    monitoring_location_id = sitelist,
    properties = c("monitoring_location_id", "parameter_code", "begin_utc", "end_utc"))

rm(sitelist, delta_sites)
```

```{r}
#filter out Delta sites that don't have potential chlorophyll-related data (parameter code 32306 is in situ fChl)
delta_chl_sites <- deltasite_metadata %>% 
    filter(parameter_code == "32316")
#lu_delta_pcodes <- unique(delta_chl_sites$parameter_code)
lu_delta_chl_sites <- unique(delta_chl_sites$monitoring_location)
```

```{r}
#Only 3 years are allowed to be pulled at a time, create list to iterate through
mindate <- min(delta_chl_sites$begin_utc, na.rm = TRUE)
mindate <- as.Date(mindate)
#mindate <- as.character(mindate)

date_limit_1 <- as.character(mindate)
date_limit_2 <- as.character(mindate + years(3))
date_limit_3 <- as.character(mindate + years(6))
date_limit_4 <- as.character(mindate + years(9))
date_limit_5 <- as.character(mindate + years(12))

maxdate <- max(delta_chl_sites$end_utc, na.rm = TRUE)
maxdate <- as.Date(maxdate)
date_limit_6 <- as.character(maxdate)

rm(mindate, maxdate)
```

```{r}
#Now grab the data (this is for daily sensor chl only in ug/l)
delta_chl_data_1 <- dataRetrieval::read_waterdata_continuous(monitoring_location_id = lu_delta_chl_sites,
                              parameter_code = "32316",
                              time = c(date_limit_1, date_limit_2)
                              )
                              
delta_chl_data_2 <- dataRetrieval::read_waterdata_continuous(monitoring_location_id = lu_delta_chl_sites,
                              parameter_code = "32316",
                              time = c(date_limit_2, date_limit_3)
                              )
                              
delta_chl_data_3 <- dataRetrieval::read_waterdata_continuous(monitoring_location_id = lu_delta_chl_sites,
                              parameter_code = "32316",
                              time = c(date_limit_3, date_limit_4)
                              )
                              
delta_chl_data_4 <- dataRetrieval::read_waterdata_continuous(monitoring_location_id = lu_delta_chl_sites,
                              parameter_code = "32316",
                              time = c(date_limit_4, date_limit_5)
                              )
                              
delta_chl_data_5 <- dataRetrieval::read_waterdata_continuous(monitoring_location_id = lu_delta_chl_sites,
                              parameter_code = "32316",
                              time = c(date_limit_5, date_limit_6)
                              )

rm(date_limit_1, date_limit_2, date_limit_3, date_limit_4, date_limit_5, date_limit_6)
```

```{r}
#stitch together all USGS CM chl data
delta_chl_data <- rbind(delta_chl_data_1, delta_chl_data_2, delta_chl_data_3, delta_chl_data_4, delta_chl_data_5)
summary(delta_chl_data)

rm(delta_chl_data_1, delta_chl_data_2, delta_chl_data_3, delta_chl_data_4, delta_chl_data_5)
```

```{r}
#Grab only daytime data and average daily and correct per the EXO bias correction formula in Richardson, 2025
delta_chl_data$hour <- as.numeric(format(delta_chl_data$time, "%H"))

delta_chl_data$TOD <- ifelse(delta_chl_data$hour %in% c(8, 9, 10, 11, 12, 13, 14, 15, 16), "keep", "remove")
delta_chl_data$date <- as.Date(delta_chl_data$time)

daytime_delta_chl_data <- delta_chl_data %>% filter(TOD == "keep")

avg_daytime_usgs_delta_chl_data <- daytime_delta_chl_data %>%
  group_by(date, monitoring_location_id)%>%
  mutate(avg_fchl = mean(value)) %>%
  mutate(fchl_corr = 1.29*avg_fchl+0.33)%>% #Richardson EXO correction (these data should already be temp corrected)
  select(date, monitoring_location_id, avg_fchl)%>%
  distinct() %>% ungroup()
```

```{r}
#add metadata
coords <- st_coordinates(deltasite_metadata$geometry)
unique_locations <- deltasite_metadata %>%
  mutate(
    longitude = coords[, "X"],
    latitude = coords[, "Y"]
  ) %>%
  select(monitoring_location_id, latitude, longitude) %>%
  distinct()

avg_daytime_usgs_delta_chl_data <- avg_daytime_usgs_delta_chl_data%>%
  left_join(
    unique_locations,
    by = "monitoring_location_id"
  )

  avg_daytime_usgs_delta_chl_data <- avg_daytime_usgs_delta_chl_data%>%
    rename(station = monitoring_location_id)

rm(coords, daytime_delta_chl_data, delta_chl_data, delta_chl_sites, deltasite_metadata, unique_locations, lu_delta_chl_sites)
```

### DWR real time monitoring
```{r}
#This script was adapted from one written by Schuyler Nardelli (USGS)

packageId <-"edi.1177.8" # Version updated 05/2026
#read_data_package_archive(packageId, path = tempdir()) 
res <- read_data_entity_names(packageId)

entityName <- "Station Metadata"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
stations <- readr::read_csv(file = raw)
stations <- stations %>% select(station, latitude, longitude, EXO2_start_date, EXO2_start_time)

stations$EXO2_start_date<- as.character(stations$EXO2_start_date)
stations$EXO2_start_time<- as.character(stations$EXO2_start_time)

stations$datetime <- paste(stations$EXO2_start_date, stations$EXO2_start_time, sep = " ")
stations <- stations %>% mutate(datetime = ymd_hms(datetime))

#indicate when data were collected by different instruments

HON_exodate <- stations %>% filter(station == "HON")
HON_exodate <- HON_exodate$datetime

MAL_exodate <- stations %>% filter(station == "MAL")
MAL_exodate <- MAL_exodate$datetime

GZL_exodate <- stations %>% filter(station == "GZL")
GZL_exodate <- GZL_exodate$datetime

RYC_exodate <- stations %>% filter(station == "RYC")
RYC_exodate <- RYC_exodate$datetime

ANH_exodate <- stations %>% filter(station == "ANH")
ANH_exodate <- ANH_exodate$datetime

FRK_exodate <- stations %>% filter(station == "FRK")
FRK_exodate <- FRK_exodate$datetime

MSD_exodate <- stations %>% filter(station == "MSD")
MSD_exodate <- MSD_exodate$datetime

PPT_exodate <- stations %>% filter(station == "PPT")
PPT_exodate <- PPT_exodate$datetime

RRI_exodate <- stations %>% filter(station == "RRI")
RRI_exodate <- RRI_exodate$datetime

RVB_exodate <- stations %>% filter(station == "RVB")
RVB_exodate <- RVB_exodate$datetime

SJR_exodate <- stations %>% filter(station == "SJR")
SJR_exodate <- SJR_exodate$datetime

SRH_exodate <- stations %>% filter(station == "SRH")
SRH_exodate <- SRH_exodate$datetime

SSI_exodate <- stations %>% filter(station == "SSI")
SSI_exodate <- SSI_exodate$datetime

TWI_exodate <- stations %>% filter(station == "TWI")
TWI_exodate <- TWI_exodate$datetime


entityName <- "ANH"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
ANH <- readr::read_csv(file = raw)
ANH <- ANH %>% rename(Chl=fluorescence)%>% 
  mutate(Date = as.character(date))%>% 
  mutate(Time = as.character(time))

ANH$datetime <- paste(ANH$Date, ANH$Time, sep = " ")

ANH <- ANH %>% 
  mutate(datetime = ymd_hms(datetime))%>% 
  mutate(instrument = if_else(datetime >= ANH_exodate, "EXO", "6600"))

entityName <- "HON"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
HON <- readr::read_csv(file = raw)
HON <- HON %>% rename(Chl=fluorescence)%>% 
  mutate(Date = as.character(date))%>% 
  mutate(Time = as.character(time))

HON$datetime <- paste(HON$Date, HON$Time, sep = " ")

HON <- HON %>% 
  mutate(datetime = ymd_hms(datetime))%>% 
  mutate(instrument = if_else(datetime >= HON_exodate, "EXO", "6600"))

entityName <- "MAL"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
MAL <- readr::read_csv(file = raw)
MAL <- MAL %>% rename(Chl=fluorescence)%>% 
  mutate(Date = as.character(date))%>% 
  mutate(Time = as.character(time))

MAL$datetime <- paste(MAL$Date, MAL$Time, sep = " ")

MAL <- MAL %>% 
  mutate(datetime = ymd_hms(datetime))%>% 
  mutate(instrument = if_else(datetime >= MAL_exodate, "EXO", "6600"))

entityName <- "GZL"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
GZL <- readr::read_csv(file = raw)
GZL <- GZL %>% rename(Chl=fluorescence)%>% 
  mutate(Date = as.character(date))%>% 
  mutate(Time = as.character(time))

GZL$datetime <- paste(GZL$Date, GZL$Time, sep = " ")

GZL <- GZL %>% 
  mutate(datetime = ymd_hms(datetime))%>% 
  mutate(instrument = if_else(datetime >= GZL_exodate, "EXO", "6600"))

entityName <- "FRK"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
FRK <- readr::read_csv(file = raw)
FRK <- FRK %>% rename(Chl=fluorescence)%>% 
  mutate(Date = as.character(date))%>% 
  mutate(Time = as.character(time))

FRK$datetime <- paste(FRK$Date, FRK$Time, sep = " ")

FRK <- FRK %>% 
  mutate(datetime = ymd_hms(datetime))%>% 
  mutate(instrument = if_else(datetime >= FRK_exodate, "EXO", "6600"))

entityName <- "RYC"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
RYC <- readr::read_csv(file = raw)
RYC <- RYC %>% rename(Chl=fluorescence)%>% 
  mutate(Date = as.character(date))%>% 
  mutate(Time = as.character(time))

RYC$datetime <- paste(RYC$Date, RYC$Time, sep = " ")

RYC <- RYC %>% 
  mutate(datetime = ymd_hms(datetime))%>% 
  mutate(instrument = if_else(datetime >= RYC_exodate, "EXO", "6600"))

entityName <- "MSD"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
MSD <- readr::read_csv(file = raw)
MSD <- MSD %>% rename(Chl=fluorescence)%>% 
  mutate(Date = as.character(date))%>% 
  mutate(Time = as.character(time))

MSD$datetime <- paste(MSD$Date, MSD$Time, sep = " ")

MSD <- MSD %>% 
  mutate(datetime = ymd_hms(datetime))%>% 
  mutate(instrument = if_else(datetime >= MSD_exodate, "EXO", "6600"))

entityName <- "PPT"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
PPT <- readr::read_csv(file = raw)
PPT <- PPT %>% rename(Chl=fluorescence)%>% 
  mutate(Date = as.character(date))%>% 
  mutate(Time = as.character(time))

PPT$datetime <- paste(PPT$Date, PPT$Time, sep = " ")

PPT <- PPT %>% 
  mutate(datetime = ymd_hms(datetime))%>% 
  mutate(instrument = if_else(datetime >= PPT_exodate, "EXO", "6600"))

entityName <- "RRI"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
RRI <- readr::read_csv(file = raw)
RRI <- RRI %>% rename(Chl=fluorescence)%>% 
  mutate(Date = as.character(date))%>% 
  mutate(Time = as.character(time))

RRI$datetime <- paste(RRI$Date, RRI$Time, sep = " ")

RRI <- RRI %>% 
  mutate(datetime = ymd_hms(datetime))%>% 
  mutate(instrument = if_else(datetime >= RRI_exodate, "EXO", "6600"))

entityName <- "RVB"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
RVB <- readr::read_csv(file = raw)
RVB <- RVB %>% rename(Chl=fluorescence)%>% 
  mutate(Date = as.character(date))%>% 
  mutate(Time = as.character(time))

RVB$datetime <- paste(RVB$Date, RVB$Time, sep = " ")

RVB <- RVB %>% 
  mutate(datetime = ymd_hms(datetime))%>% 
  mutate(instrument = if_else(datetime >= RVB_exodate, "EXO", "6600"))

entityName <- "SJR"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
SJR <- readr::read_csv(file = raw)
SJR <- SJR %>% rename(Chl=fluorescence)%>% 
  mutate(Date = as.character(date))%>% 
  mutate(Time = as.character(time))

SJR$datetime <- paste(SJR$Date, SJR$Time, sep = " ")

SJR <- SJR %>% 
  mutate(datetime = ymd_hms(datetime))%>% 
  mutate(instrument = if_else(datetime >= SJR_exodate, "EXO", "6600"))

entityName <- "SRH"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
SRH <- readr::read_csv(file = raw)
SRH <- SRH %>% rename(Chl=fluorescence)%>% 
  mutate(Date = as.character(date))%>% 
  mutate(Time = as.character(time))

SRH$datetime <- paste(SRH$Date, SRH$Time, sep = " ")

SRH <- SRH %>% 
  mutate(datetime = ymd_hms(datetime))%>% 
  mutate(instrument = if_else(datetime >= SRH_exodate, "EXO", "6600"))

entityName <- "SSI"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
SSI <- readr::read_csv(file = raw)
SSI <- SSI %>% rename(Chl=fluorescence)%>% 
  mutate(Date = as.character(date))%>% 
  mutate(Time = as.character(time))

SSI$datetime <- paste(SSI$Date, SSI$Time, sep = " ")

SSI <- SSI %>% 
  mutate(datetime = ymd_hms(datetime))%>% 
  mutate(instrument = if_else(datetime >= SSI_exodate, "EXO", "6600"))

entityName <- "TWI"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
TWI <- readr::read_csv(file = raw)
TWI <- TWI %>% rename(Chl=fluorescence)%>% 
  mutate(Date = as.character(date))%>% 
  mutate(Time = as.character(time))

TWI$datetime <- paste(TWI$Date, TWI$Time, sep = " ")

TWI <- TWI %>% 
  mutate(datetime = ymd_hms(datetime))%>% 
  mutate(instrument = if_else(datetime >= TWI_exodate, "EXO", "6600"))


rm(entityName,entityId,res,raw,packageId)
# Merge all stations ------------------------------------------------------

CM_DWR <- rbind(HON,MAL,GZL,RYC,ANH,FRK,MSD,PPT,RRI,RVB,SJR,SRH,SSI,TWI)

rm(HON,MAL,GZL,RYC,ANH,FRK,MSD,PPT,RRI,RVB,SJR,SRH,SSI,TWI)
rm(HON_exodate,MAL_exodate,GZL_exodate,RYC_exodate,ANH_exodate,FRK_exodate,MSD_exodate,PPT_exodate,RRI_exodate,RVB_exodate,SJR_exodate,SRH_exodate,SSI_exodate,TWI_exodate)
```

```{r}
#correct EXO chl data and make daytime averages
 CM_DWR <- CM_DWR %>% 
  mutate(hour = hour(time))

CM_DWR$TOD <- ifelse(CM_DWR$hour %in% c(8, 9, 10, 11, 12, 13, 14, 15, 16), "keep", "remove")

CM_DWR <- CM_DWR %>% filter(TOD == "keep")

avg_daytime_dwr_delta_chl_data <- CM_DWR %>%
  mutate(chl_tempcorr = Chl/(1+(0.01*(watertemperature-25))))%>% #Watras temerature correction
  mutate(fchl_corr = if_else(instrument == "EXO", 1.29*chl_tempcorr+0.33, chl_tempcorr))%>% #Richardson EXO correction
  group_by(date, station, instrument)%>%
  mutate(avg_fchl = mean(fchl_corr)) %>%
  select(date, station, avg_fchl, instrument)%>%
  filter(!is.na(avg_fchl))%>% 
  distinct()%>% 
  ungroup() 


summary(avg_daytime_dwr_delta_chl_data)
```

```{r}
# add metadata
stations <-  stations %>% 
  select(station, latitude, longitude)

avg_daytime_dwr_delta_chl_data <- avg_daytime_dwr_delta_chl_data %>%
  left_join(
    stations,
    by = "station"
  )
  rm(CM_DWR, stations)
```

### DWR NCRO real time monitoring
```{r}
packageId <-"edi.2180.1" # Version updated 10/2025
res <- read_data_entity_names(packageId)

entityName <- "Stations Metadata"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
stations <- readr::read_csv(file = raw)
stations <- stations %>% select(Station_Code, Latitude, Longitude) %>%
    rename(station = Station_Code, latitude = Latitude, longitude = Longitude)

entityName <- "Instrument Type Date Ranges"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
instruments <- readr::read_csv(file = raw)
instruments$Start_Time_YSIEXO2 <- as.character(instruments$Start_Time_YSIEXO2)

instruments$datetime <- paste(instruments$Start_Date_YSIEXO2, instruments$Start_Time_YSIEXO2, sep = " ")
instruments <- instruments %>% mutate(datetime = mdy_hms(datetime))

#indicate when data were collected by different instruments

GLC_exodate <- instruments %>% filter(Station_Code == "GLC")
GLC_exodate <- GLC_exodate$datetime

GLE_exodate <- instruments %>% filter(Station_Code == "GLE")
GLE_exodate <- GLE_exodate$datetime

MHO_exodate <- instruments %>% filter(Station_Code == "MHO")
MHO_exodate <- MHO_exodate$datetime

MRU_exodate <- instruments %>% filter(Station_Code == "MRU")
MRU_exodate <- MRU_exodate$datetime

MRX_exodate <- instruments %>% filter(Station_Code == "MRX")
MRX_exodate <- MRX_exodate$datetime

ODM_exodate <- instruments %>% filter(Station_Code == "ODM")
ODM_exodate <- ODM_exodate$datetime

OH1_exodate <- instruments %>% filter(Station_Code == "OH1")
OH1_exodate <- OH1_exodate$datetime

ORI_exodate <- instruments %>% filter(Station_Code == "ORI")
ORI_exodate <- ORI_exodate$datetime

ORM_exodate <- instruments %>% filter(Station_Code == "ORM")
ORM_exodate <- ORM_exodate$datetime

ORX_exodate <- instruments %>% filter(Station_Code == "ORX")
ORX_exodate <- ORX_exodate$datetime

TWA_exodate <- instruments %>% filter(Station_Code == "TWA")
TWA_exodate <- TWA_exodate$datetime

WCI_exodate <- instruments %>% filter(Station_Code == "WCI")
WCI_exodate <- WCI_exodate$datetime

VCU_exodate <- instruments %>% filter(Station_Code == "VCU")
VCU_exodate <- VCU_exodate$datetime


entityName <- "GLC"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
GLC <- readr::read_csv(file = raw)
GLC <- GLC %>% rename(station=Station_Code)  %>%
  mutate(Date_Time = mdy_hms(Date_Time)) %>%
  mutate(instrument = if_else(Date_Time >= GLC_exodate, "EXO", "6600"))

entityName <- "GLE"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
GLE<- readr::read_csv(file = raw)
GLE <- GLE %>% rename(station=Station_Code)  %>%
  mutate(Date_Time = mdy_hms(Date_Time)) %>%
  mutate(instrument = if_else(Date_Time >= GLE_exodate, "EXO", "6600"))

entityName <- "MHO"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
MHO <- readr::read_csv(file = raw)
MHO<- MHO %>% rename(station=Station_Code)  %>%
  mutate(Date_Time = mdy_hms(Date_Time)) %>%
  mutate(instrument = if_else(Date_Time >= MHO_exodate, "EXO", "6600"))

entityName <- "MRU"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
MRU<- readr::read_csv(file = raw)
MRU<- MRU %>% rename(station=Station_Code)  %>%
  mutate(Date_Time = mdy_hms(Date_Time)) %>%
  mutate(instrument = if_else(Date_Time >= MRU_exodate, "EXO", "6600"))

entityName <- "MRX"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
MRX<- readr::read_csv(file = raw)
MRX<- MRX %>% rename(station=Station_Code)  %>%
  mutate(Date_Time = mdy_hms(Date_Time)) %>%
  mutate(instrument = if_else(Date_Time >= MRX_exodate, "EXO", "6600"))

entityName <- "ODM"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
ODM<- readr::read_csv(file = raw)
ODM<- ODM %>% rename(station=Station_Code)  %>%
  mutate(Date_Time = mdy_hms(Date_Time)) %>%
  mutate(instrument = if_else(Date_Time >= ODM_exodate, "EXO", "6600"))

entityName <- "OH1"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
OH1<- readr::read_csv(file = raw)
OH1<- OH1 %>% rename(station=Station_Code)  %>%
  mutate(Date_Time = mdy_hms(Date_Time)) %>%
  mutate(instrument = if_else(Date_Time >= OH1_exodate, "EXO", "6600"))

entityName <- "ORI"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
ORI<- readr::read_csv(file = raw)
ORI<- ORI %>% rename(station=Station_Code)  %>%
  mutate(Date_Time = mdy_hms(Date_Time)) %>%
  mutate(instrument = if_else(Date_Time >= ORI_exodate, "EXO", "6600"))

entityName <- "ORM"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
ORM<- readr::read_csv(file = raw)
ORM<- ORM %>% rename(station=Station_Code)  %>%
  mutate(Date_Time = mdy_hms(Date_Time)) %>%
  mutate(instrument = if_else(Date_Time >= ORM_exodate, "EXO", "6600"))

entityName <- "ORX"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
ORX<- readr::read_csv(file = raw)
ORX<- ORX %>% rename(station=Station_Code)  %>%
  mutate(Date_Time = mdy_hms(Date_Time)) %>%
  mutate(instrument = if_else(Date_Time >= ORX_exodate, "EXO", "6600"))

entityName <- "TWA"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
TWA <- readr::read_csv(file = raw)
TWA<- TWA %>% rename(station=Station_Code)  %>%
  mutate(Date_Time = mdy_hms(Date_Time)) %>%
  mutate(instrument = if_else(Date_Time >= TWA_exodate, "EXO", "6600"))

entityName <- "WCI"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
WCI<- readr::read_csv(file = raw)
WCI<- WCI %>% rename(station=Station_Code)  %>%
  mutate(Date_Time = mdy_hms(Date_Time)) %>%
  mutate(instrument = if_else(Date_Time >= WCI_exodate, "EXO", "6600"))

entityName <- "VCU"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
VCU<- readr::read_csv(file = raw)
VCU<- VCU %>% rename(station=Station_Code)  %>%
  mutate(Date_Time = mdy_hms(Date_Time)) %>%
  mutate(instrument = if_else(Date_Time >= VCU_exodate, "EXO", "6600"))

cm_ncro<- rbind(GLC, GLE, MHO, MRU, MRX, ODM, OH1, ORI, ORM, ORX, TWA, WCI, VCU)
rm(GLC_exodate, GLE_exodate, MHO_exodate, MRU_exodate, MRX_exodate, ODM_exodate, OH1_exodate, ORI_exodate, ORM_exodate, ORX_exodate, TWA_exodate, WCI_exodate, VCU_exodate)
rm(GLC, GLE, MHO, MRU, MRX, ODM, OH1, ORI, ORM, ORX, TWA, WCI, VCU, res, raw, entityName, entityId, packageId, instruments)
```

```{r}
# calculate daytime averages and add metadata
cm_ncro <- cm_ncro %>% 
   rename(datetime =Date_Time)%>% 
    mutate(hour = hour(datetime))

cm_ncro$TOD <- ifelse(cm_ncro$hour %in% c(8, 9, 10, 11, 12, 13, 14, 15, 16), "keep", "remove")

cm_ncro <- cm_ncro %>% filter(TOD == "keep")

avg_daytime_ncro_delta_chl_data <- cm_ncro %>%
    mutate(date = as_date(datetime))%>%
    mutate(chl_tempcorr = Chlorophyll/(1+(0.01*(Water_Temperature-25))))%>% #Watras temerature correction
    mutate(fchl_corr = if_else(instrument == "EXO", 1.29*chl_tempcorr+0.33, Chlorophyll))%>% #Richardson EXO correction
    group_by(date, station, instrument)%>%
    mutate(avg_fchl = mean(fchl_corr)) %>%
    select(date, station, avg_fchl, instrument)%>%
    filter(!is.na(avg_fchl))%>% 
    distinct()%>% 
    ungroup() 

summary(avg_daytime_ncro_delta_chl_data)
```

```{r}
avg_daytime_ncro_delta_chl_data  <- avg_daytime_ncro_delta_chl_data  %>%
  left_join(
    stations,
    by = "station"
  )

rm(cm_ncro, stations)
```

## Combine all continuous monitoring chlorophyll data

```{r}
avg_daytime_usgs_delta_chl_data$instrument <- "EXO"

avg_daytime_dwr_delta_chl_data$group <- "dwr_rtm"
avg_daytime_ncro_delta_chl_data$group <- "dwr_ncro"
avg_daytime_usgs_delta_chl_data$group <- "usgs"
cm_daytime_avg_chl <- rbind(avg_daytime_dwr_delta_chl_data, avg_daytime_ncro_delta_chl_data, avg_daytime_usgs_delta_chl_data)

rm(avg_daytime_dwr_delta_chl_data, avg_daytime_ncro_delta_chl_data, avg_daytime_usgs_delta_chl_data)

#write.csv(cm_daytime_avg_chl, file = "data/01_processed/01_chl_cm_daytime_avg.csv")
```

## Shaela's Workspace (Emily jumping in too :) 

### Discrete WQ Data Package

#### Install, load package and read in data


```{r}
#devtools::install_github("InteragencyEcologicalProgram/discretewq")

library(discretewq)
library(dplyr)
library(sf)
```
```{r}
discrete_raw <- wq(
  Sources = c(
    "20mm",
    "Baystudy",
    "DJFMP",
    "DOP",
    "EDSM",
    "EMP",
    "FMWT",
    "NCRO",
    "SDO",
    "SKT",
    "SLS",
    "STN",
    "Suisun",
    "USBR",
    "USGS_CAWSC",
    "USGS_SFBS",
    "YBFMP"
  )
)

str(discrete_raw)
```

#### Data exploration
```{r}
#filter to chlorophyll
discrete_chla <- discrete_raw %>% 
    filter(!is.na(Chlorophyll)) %>% filter(!is.na(Latitude)) %>%
    select(Source, Station, Latitude, Longitude, Date, Datetime, Depth, Chlorophyll, Chlorophyll_Sign) %>%
    rename(group = Source,
          station = Station,
          chla = Chlorophyll,
          chla_rmk = Chlorophyll_Sign,
          depth = Depth, 
          latitude = Latitude,
          longitude = Longitude,
          date = Date,
          datetime = Datetime) %>%
    mutate(sample_type = "discrete",
          date = as.Date(date))

#see which surveys have chlorophyll data
discrete_chla %>% 
    distinct(group)

#explore what the chlorophyll notes are 
unique(discrete_chla$chla_rmk)
```

```{r}
#save data to 01 folder
write.csv(discrete_chla, file = "data/01_processed/01_chl_iep.csv")
```

```{r}
#convert to a spatial object
discrete_chla_sf <- discrete_chla %>% 
    st_as_sf(coords = c('Longitude', 'Latitude'),
             crs = 4326,
             remove = FALSE)

test <- discrete_chla_sf %>% 
    filter(Source == "DOP") %>% 
    distinct(Latitude, Longitude, .keep_all = TRUE)
```



## Spencer's Workspace

### e.g., name of data source

## Running spatial data packages
```{r}
### Data wrangling packages:
library(readr)
library(dplyr)
library(tidyr)
library(here)

### Simple Features vector spatial data:
library(sf)

### Data visualization packages:
library(ggplot2)
library(leaflet)
library(ggspatial) 
library(prettymapr)
library(EDIutils)
### ggspatial depends on but does not import prettymapr for internal functions
```

```{r}
packageId <-"edi.1537.4" 
read_data_package_archive(packageId, path = tempdir()) 
res <- read_data_entity_names(packageId)

entityName <- "Chlorophyll and Covariate Data"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
chla_special_sf <- readr::read_csv(file = raw)
```

```{r}
packageId <-"edi.1537.4" 
#read_data_package_archive(packageId, path = tempdir()) 
res <- read_data_entity_names(packageId)

entityName <- "Station locations"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
chla_stations <- readr::read_csv(file = raw)
```
```{r}
chla_special_sf <-read_csv(here("data/model_chla_covars.csv"))

chla_stations <- read_csv(here("data/stations.csv"))
```

```{r}
chla_lat_long <- chla_stations %>%
    select(station, latitude, longitude) %>%
    mutate(latitude = round(latitude, 3), 
        longitude = round(longitude, 3)    
        ) %>%
    distinct()
chla_joined <- chla_special_sf %>%
    full_join(chla_lat_long, by = "station")

write.csv(chla_joined, "data/01_tidy_model_chla_covars.csv")
```


```{r}
chla_4326_sf <- st_as_sf(chla_joined,
                        coords = c('longitude', 'latitude'),
                        crs = 4326)
```


```{r}
plot(chla_4326_sf %>%
    select(chlorophyll))
```


## Making a Map with the Chlorophyll Stations

```{r}
ggplot(chla_4326_sf) +
    geom_sf(aes(fill = chlorophyll)) +
    labs(fill = "chlorophyll") +
    scale_fill_continuous(low = "blue", high =  "green") +
    theme_bw()
```


```{r}
chla_3857_sf <- st_transform(chla_4326_sf, 
                            crs = 3857)
```


```{r}
plot(chla_3857_sf %>%
        select(chlorophyll))
 ggplot(chla_3857_sf) +
     ggspatial::annotation_map_tile(type = "osm", zoom = 10, progress = 'none') +
       geom_sf(aes(fill = chlorophyll)) +
     labs(fill = "chlorophyll") +
     scale_fill_continuous(low = "blue", high =  "green") +
     theme_bw()
```


```{r}
packageId <-"edi.1653.1" 
read_data_package_archive(packageId, path = tempdir()) 
res <- read_data_entity_names(packageId)

entityName <- "Chlorophyll_1975_2021"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
chla_drought <- readr::read_csv(file = raw)
```
