---
title: "Grabbing Chlorophyll data"
format: html: embed-resources: TRUE
author: Emily Richardson, Shaela Noble, Spencer Fern (others tbd)

---
# This document shows steps that the 2026 NCEAS CHAB team is using to harmonize chlorophyll data

## Emily's Workspace

```{r}
#load packages
library(dataRetrieval)
library(dplyr)
library(purrr)
library(sf)
library(stringr)
library(DT)
library(lubridate)
library(DBI)
library(EDIutils)
```

### ScienceBase
#### Used https://www.sciencebase.gov/catalog/items/get?q=chlorophyll%20sacramento to identify 22 datasets (links in excel file in Box, will downloaded individually and added to raw data folder)




### Water data for the nation (USGS)
```{r}
# Identify Delta surface water sites
out_sf <- 
  dataRetrieval::read_waterdata_monitoring_location(
    state_name = "California", 
    hydrologic_unit_code = "18020163",
    agency_name = "U.S. Geological Survey",
    properties = c("monitoring_location_id", "site_type", "hydrologic_unit_code", "county_name", "agency_name"))

sites <- out_sf %>%
  mutate(lon = st_coordinates(.)[,1],
         lat = st_coordinates(.)[,2])

delta_sites <- sites %>% filter((between(lat, 37.585545, 38.566034)) &  between(lon, -122.292649, -121.131962))

rm(out_sf, sites)
```

```{r}
sitelist <- delta_sites$monitoring_location_id
deltasite_metadata <- read_waterdata_ts_meta(
    monitoring_location_id = sitelist,
    properties = c("monitoring_location_id", "parameter_code", "begin_utc", "end_utc"))

rm(sitelist, delta_sites)
```

```{r}
#filter out Delta sites that don't have potential chlorophyll-related data (parameter code 32306 is in situ fChl)
delta_chl_sites <- deltasite_metadata %>% 
    filter(parameter_code == "32316")
#lu_delta_pcodes <- unique(delta_chl_sites$parameter_code)
lu_delta_chl_sites <- unique(delta_chl_sites$monitoring_location)
```

```{r}
#Only 3 years are allowed to be pulled at a time, create list to iterate through
mindate <- min(delta_chl_sites$begin_utc, na.rm = TRUE)
mindate <- as.Date(mindate)
#mindate <- as.character(mindate)

date_limit_1 <- as.character(mindate)
date_limit_2 <- as.character(mindate + years(3))
date_limit_3 <- as.character(mindate + years(6))
date_limit_4 <- as.character(mindate + years(9))
date_limit_5 <- as.character(mindate + years(12))

maxdate <- max(delta_chl_sites$end_utc, na.rm = TRUE)
maxdate <- as.Date(maxdate)
date_limit_6 <- as.character(maxdate)

rm(mindate, maxdate)
```

```{r}
#Now grab the data (this is fordaily sensor chl only in ug/l)
delta_chl_data_1 <- dataRetrieval::read_waterdata_continuous(monitoring_location_id = lu_delta_chl_sites,
                              parameter_code = "32316",
                              time = c(date_limit_1, date_limit_2)
                              )
                              
delta_chl_data_2 <- dataRetrieval::read_waterdata_continuous(monitoring_location_id = lu_delta_chl_sites,
                              parameter_code = "32316",
                              time = c(date_limit_2, date_limit_3)
                              )
                              
delta_chl_data_3 <- dataRetrieval::read_waterdata_continuous(monitoring_location_id = lu_delta_chl_sites,
                              parameter_code = "32316",
                              time = c(date_limit_3, date_limit_4)
                              )
                              
delta_chl_data_4 <- dataRetrieval::read_waterdata_continuous(monitoring_location_id = lu_delta_chl_sites,
                              parameter_code = "32316",
                              time = c(date_limit_4, date_limit_5)
                              )
                              
delta_chl_data_5 <- dataRetrieval::read_waterdata_continuous(monitoring_location_id = lu_delta_chl_sites,
                              parameter_code = "32316",
                              time = c(date_limit_5, date_limit_6)
                              )

rm(date_limit_1, date_limit_2, date_limit_3, date_limit_4, date_limit_5, date_limit_6)
```

```{r}
#stitch together all USGS CM chl data
delta_chl_data <- rbind(delta_chl_data_1, delta_chl_data_2, delta_chl_data_3, delta_chl_data_4, delta_chl_data_5)
summary(delta_chl_data)

rm(delta_chl_data_1, delta_chl_data_2, delta_chl_data_3, delta_chl_data_4, delta_chl_data_5)
```

```{r}
#Grab only daytime data and average daily and correct per the EXO bias correction formula in Richardson, 2025
delta_chl_data$hour <- as.numeric(format(delta_chl_data$time, "%H"))

delta_chl_data$TOD <- ifelse(delta_chl_data$hour %in% c(8, 9, 10, 11, 12, 13, 14, 15, 16), "keep", "remove")
delta_chl_data$date <- as.Date(delta_chl_data$time)

daytime_delta_chl_data <- delta_chl_data %>% filter(TOD == "keep")

avg_daytime_usgs_delta_chl_data <- daytime_delta_chl_data %>%
  group_by(date, monitoring_location_id)%>%
  mutate(avg_fchl = mean(value)) %>%
  mutate(fchl_corr = 1.29*avg_fchl+0.33)%>%
  select(date, monitoring_location_id, fchl_corr, avg_fchl)%>%
  distinct() %>% ungroup()
```

```{r}
#add metadata
coords <- st_coordinates(deltasite_metadata$geometry)
unique_locations <- deltasite_metadata %>%
  mutate(
    longitude = coords[, "X"],
    latitude = coords[, "Y"]
  ) %>%
  select(monitoring_location_id, latitude, longitude) %>%
  distinct()

avg_daytime_usgs_delta_chl_data <- avg_daytime_usgs_delta_chl_data%>%
  left_join(
    unique_locations,
    by = "monitoring_location_id"
  )

  avg_daytime_usgs_delta_chl_data <- avg_daytime_usgs_delta_chl_data%>%
    rename(station = monitoring_location_id)

rm(coords, daytime_delta_chl_data, delta_chl_data, delta_chl_sites, deltasite_metadata, unique_locations, lu_delta_chl_sites)
```

### DWR real time monitoring
```{r}
#This script was adapted from one written by Schuyler Nardelli (USGS)

packageId <-"edi.1177.8" # Version updated 05/2026
#read_data_package_archive(packageId, path = tempdir()) 
res <- read_data_entity_names(packageId)

entityName <- "Station Metadata"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
stations <- readr::read_csv(file = raw)
stations <- stations %>% select(station, latitude, longitude)


entityName <- "ANH"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
ANH <- readr::read_csv(file = raw)
ANH <- ANH %>% rename(Station=station,Date=date,Time=time,Oxy=dissolvedoxygen,Chl=fluorescence,pH=ph,SpecCond=spc,Turb=turbidity,Temp=watertemperature)

entityName <- "FRK"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
FRK <- readr::read_csv(file = raw)
FRK <- FRK %>% rename(Station=station,Date=date,Time=time,Oxy=dissolvedoxygen,Chl=fluorescence,pH=ph,SpecCond=spc,Turb=turbidity,Temp=watertemperature)

entityName <- "GZL"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
GZL <- readr::read_csv(file = raw)
GZL <- GZL %>% rename(Station=station,Date=date,Time=time,Oxy=dissolvedoxygen,Chl=fluorescence,pH=ph,SpecCond=spc,Turb=turbidity,Temp=watertemperature)

entityName <- "HON"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
HON <- readr::read_csv(file = raw)
HON <- HON %>% rename(Station=station,Date=date,Time=time,Oxy=dissolvedoxygen,Chl=fluorescence,pH=ph,SpecCond=spc,Turb=turbidity,Temp=watertemperature)

entityName <- "MAL"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
MAL <- readr::read_csv(file = raw)
MAL <- MAL %>% rename(Station=station,Date=date,Time=time,Oxy=dissolvedoxygen,Chl=fluorescence,pH=ph,SpecCond=spc,Turb=turbidity,Temp=watertemperature)

entityName <- "MSD"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
MSD <- readr::read_csv(file = raw)
MSD <- MSD %>% rename(Station=station,Date=date,Time=time,Oxy=dissolvedoxygen,Chl=fluorescence,pH=ph,SpecCond=spc,Turb=turbidity,Temp=watertemperature)

entityName <- "PPT"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
PPT <- readr::read_csv(file = raw)
PPT <- PPT %>% rename(Station=station,Date=date,Time=time,Oxy=dissolvedoxygen,Chl=fluorescence,pH=ph,SpecCond=spc,Turb=turbidity,Temp=watertemperature)

entityName <- "RRI"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
RRI <- readr::read_csv(file = raw)
RRI <- RRI %>% rename(Station=station,Date=date,Time=time,Oxy=dissolvedoxygen,Chl=fluorescence,pH=ph,SpecCond=spc,Turb=turbidity,Temp=watertemperature)

entityName <- "RVB"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
RVB <- readr::read_csv(file = raw)
RVB <- RVB %>% rename(Station=station,Date=date,Time=time,Oxy=dissolvedoxygen,Chl=fluorescence,pH=ph,SpecCond=spc,Turb=turbidity,Temp=watertemperature)

entityName <- "RYC"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
RYC <- readr::read_csv(file = raw)
RYC <- RYC %>% rename(Station=station,Date=date,Time=time,Oxy=dissolvedoxygen,Chl=fluorescence,pH=ph,SpecCond=spc,Turb=turbidity,Temp=watertemperature)

entityName <- "SJR"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
SJR <- readr::read_csv(file = raw)
SJR <- SJR %>% rename(Station=station,Date=date,Time=time,Oxy=dissolvedoxygen,Chl=fluorescence,pH=ph,SpecCond=spc,Turb=turbidity,Temp=watertemperature)

entityName <- "SRH"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
SRH <- readr::read_csv(file = raw)
SRH <- SRH %>% rename(Station=station,Date=date,Time=time,Oxy=dissolvedoxygen,Chl=fluorescence,pH=ph,SpecCond=spc,Turb=turbidity,Temp=watertemperature)

entityName <- "SSI"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
SSI <- readr::read_csv(file = raw)
SSI <- SSI %>% rename(Station=station,Date=date,Time=time,Oxy=dissolvedoxygen,Chl=fluorescence,pH=ph,SpecCond=spc,Turb=turbidity,Temp=watertemperature)

entityName <- "TWI"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
TWI <- readr::read_csv(file = raw)
TWI <- TWI %>% rename(Station=station,Date=date,Time=time,Oxy=dissolvedoxygen,Chl=fluorescence,pH=ph,SpecCond=spc,Turb=turbidity,Temp=watertemperature)


rm(entityName,entityId,res,raw,packageId)
# Merge all stations ------------------------------------------------------

CM_DWR <- rbind(HON,MAL,GZL,RYC,ANH,FRK,MSD,PPT,RRI,RVB,SJR,SRH,SSI,TWI)

rm(HON,MAL,GZL,RYC,ANH,FRK,MSD,PPT,RRI,RVB,SJR,SRH,SSI,TWI)
```

```{r}
#correct EXO chl data and make daytime averages
 CM_DWR <- CM_DWR %>% mutate(hour = hour(Time))

CM_DWR$TOD <- ifelse(CM_DWR$hour %in% c(8, 9, 10, 11, 12, 13, 14, 15, 16), "keep", "remove")

CM_DWR <- CM_DWR %>% filter(TOD == "keep")

avg_daytime_dwr_delta_chl_data <- CM_DWR %>%
  group_by(Date, Station)%>%
  mutate(avg_fchl = mean(Chl)) %>%
  mutate(fchl_corr = 1.29*avg_fchl+0.33)%>%
  select(Date, Station, fchl_corr, avg_fchl)%>%
  rename(date = Date, station = Station)
  distinct()%>% ungroup()
```

```{r}
# add metadata
avg_daytime_dwr_delta_chl_data <- avg_daytime_dwr_delta_chl_data %>%
  left_join(
    stations,
    by = "station"
  )
  rm(CM_DWR, stations)
```

### DWR NCRO real time monitoring
```{r}
packageId <-"edi.2180.1" # Version updated 010/2025
res <- read_data_entity_names(packageId)

entityName <- "Stations Metadata"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
stations <- readr::read_csv(file = raw)
stations <- stations %>% select(Station_Code, Latitude, Longitude) %>%
    rename(station = Station_Code, latitude = Latitude, longitude = Longitude)


entityName <- "GLC"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
GLC <- readr::read_csv(file = raw)
GLC <- GLC %>% rename(station=Station_Code)

entityName <- "GLE"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
GLE<- readr::read_csv(file = raw)
GLE <- GLE %>% rename(station=Station_Code)

entityName <- "MHO"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
MHO <- readr::read_csv(file = raw)
MHO <- MHO %>% rename(station=Station_Code)

entityName <- "MRU"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
MRU<- readr::read_csv(file = raw)
MRU <- MRU%>% rename(station=Station_Code)

entityName <- "MRX"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
MRX<- readr::read_csv(file = raw)
MRX <- MRX %>% rename(station=Station_Code)

entityName <- "ODM"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
ODM<- readr::read_csv(file = raw)
ODM <- ODM %>% rename(station=Station_Code)

entityName <- "OH1"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
OH1<- readr::read_csv(file = raw)
OH1 <- OH1 %>% rename(station=Station_Code)

entityName <- "ORI"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
ORI<- readr::read_csv(file = raw)
ORI <-ORI %>% rename(station=Station_Code)

entityName <- "ORM"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
ORM<- readr::read_csv(file = raw)
ORM <- ORM %>% rename(station=Station_Code)

entityName <- "ORX"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
ORX<- readr::read_csv(file = raw)
ORX <- ORX %>% rename(station=Station_Code)

entityName <- "TWA"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
TWA <- readr::read_csv(file = raw)
TWA <- TWA %>% rename(station=Station_Code)

entityName <- "WCI"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
WCI<- readr::read_csv(file = raw)
WCI <-WCI%>% rename(station=Station_Code)

entityName <- "VCU"; entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
VCU<- readr::read_csv(file = raw)
VCU <- VCU %>% rename(station=Station_Code)

cm_ncro<- rbind(GLC, GLE, MHO, MRU, MRX, ODM, OH1, ORI, ORM, ORX, TWA, WCI, VCU)

rm(GLC, GLE, MHO, MRU, MRX, ODM, OH1, ORI, ORM, ORX, TWA, WCI, VCU, res, raw, entityName, entityId, packageId)
```

```{r}
# calculate daytime averages and add metadata
cm_ncro <- cm_ncro %>% 
    mutate(datetime = mdy_hms(Date_Time))%>% 
    mutate(hour = hour(datetime))

cm_ncro$TOD <- ifelse(cm_ncro$hour %in% c(8, 9, 10, 11, 12, 13, 14, 15, 16), "keep", "remove")

cm_ncro <- cm_ncro %>% filter(TOD == "keep")

avg_daytime_ncro_delta_chl_data <- cm_ncro %>%
    mutate(date = as_date(datetime))%>%
  group_by(date, station)%>%
  mutate(avg_fchl = mean(Chlorophyll)) %>%
  mutate(fchl_corr = 1.29*avg_fchl+0.33)%>%
  select(date, station, fchl_corr, avg_fchl)%>%
  distinct()%>% ungroup()

avg_daytime_ncro_delta_chl_data  <- avg_daytime_ncro_delta_chl_data  %>%
  left_join(
    stations,
    by = "station"
  )

rm(cm_ncro, stations)
```

## Combine all continuous monitoring chlorophyll data

```{r}
avg_daytime_dwr_delta_chl_data$group <- "dwr_rtm"
avg_daytime_ncro_delta_chl_data$group <- "dwr_ncro"
avg_daytime_usgs_delta_chl_data$group <- "usgs"
cm_daytime_avg_chl <- rbind(avg_daytime_dwr_delta_chl_data, avg_daytime_ncro_delta_chl_data, avg_daytime_usgs_delta_chl_data)

rm(avg_daytime_dwr_delta_chl_data, avg_daytime_ncro_delta_chl_data, avg_daytime_usgs_delta_chl_data)
```

## Shaela's Workspace

### Discrete WQ Data Package

#### Install, load package and read in data


```{r}
#devtools::install_github("InteragencyEcologicalProgram/discretewq")

library(discretewq)
library(dplyr)
library(sf)
```
```{r}
discrete_raw <- wq(
  Sources = c(
    "20mm",
    "Baystudy",
    "DJFMP",
    "DOP",
    "EDSM",
    "EMP",
    "FMWT",
    "NCRO",
    "SDO",
    "SKT",
    "SLS",
    "STN",
    "Suisun",
    "USBR",
    "USGS_CAWSC",
    "USGS_SFBS",
    "YBFMP"
  )
)

str(discrete_raw)
```

#### Data exploration
```{r}
#filter to chlorophyll
discrete_chla <- discrete_raw %>% 
    filter(!is.na(Chlorophyll))

#see which surveys have chlorophyll data
discrete_chla %>% 
    distinct(Source)
```

```{r}
#filter out any NA in coordinates columns and convert to a spatial object
discrete_chla_sf <- discrete_chla %>% 
    filter(!is.na(Latitude)) %>% 
    st_as_sf(coords = c('Longitude', 'Latitude'),
             crs = 4326,
             remove = FALSE)

test <- discrete_chla_sf %>% 
    filter(Source == "DOP") %>% 
    distinct(Latitude, Longitude, .keep_all = TRUE)
```



## Spencer's Workspace

### e.g., name of data source

## Running spatial data packages
```{r}
### Data wrangling packages:
library(readr)
library(dplyr)
library(tidyr)
library(here)

### Simple Features vector spatial data:
library(sf)

### Data visualization packages:
library(ggplot2)
library(leaflet)
library(ggspatial) 
library(prettymapr) ### ggspatial depends on but does not import prettymapr for internal functions
```

```{r}
chla_special_sf <-read_csv(here("data/model_chla_covars.csv"))

chla_stations <- read_csv(here("data/stations.csv"))
```

```{r}
chla_lat_long <- chla_stations %>%
    select(station, latitude, longitude) %>%
    mutate(latitude = round(latitude, 3), 
        longitude = round(longitude, 3)    
        ) %>%
    distinct()
chla_joined <- chla_special_sf %>%
    full_join(chla_lat_long, by = "station")

write.csv(chla_joined, "data/01_tidy_model_chla_covars.csv")
```


```{r}
chla_4326_sf <- st_as_sf(chla_joined,
                        coords = c('longitude', 'latitude'),
                        crs = 4326)
```


```{r}
plot(chla_4326_sf %>%
    select(chlorophyll))
```


## Making a Map with the Chlorophyll Stations

```{r}
ggplot(chla_4326_sf) +
    geom_sf(aes(fill = chlorophyll)) +
    labs(fill = "chlorophyll") +
    scale_fill_continuous(low = "blue", high =  "green") +
    theme_bw()
```


```{r}
chla_3857_sf <- st_transform(chla_4326_sf, 
                            crs = 3857)
```


```{r}
plot(chla_3857_sf %>%
        select(chlorophyll))
 ggplot(chla_3857_sf) +
     ggspatial::annotation_map_tile(type = "osm", zoom = 10, progress = 'none') +
       geom_sf(aes(fill = chlorophyll)) +
     labs(fill = "chlorophyll") +
     scale_fill_continuous(low = "blue", high =  "green") +
     theme_bw()
```
