---
title: "Download and wrangle hydro datasets"
format: 
  html:
    toc: true
    number-sections: true
    embed-resources: true
authors: Michelle Stern and Nick Rowlands
---

# Brainstorm

The primary environmental factors shaping phytoplankton abundance and community composition are salinity, hydrodynamics, stratification, light, temperature, nutrients, and time. (https://essopenarchive.org/doi/pdf/10.22541/essoar.176495220.03206867/v1) \* Positive: high nutrients, low turbulence, long residence time, adequate light, warm temperatures, low grazing rates, seed population/refugia, low biodiversity. \* Modulating Factors: strong environmental gradients like stratification and salinity, nutrient frms and ratios, diverse habitats/connectivity, selective grazing, predator/prey co-evolution, adequate light, toxin production \* Negative: low nutrients\*, high turbulence, short residence time, low light, high grazing rates, high biodiversity, extreme temperatures warm or cold

Variables from SCHISM: water temperature, flow, turbidity, salinity, residence time (average for a certain month? season?) how to get an index for stratification, light? Variables from climate datasets: wind, solar radiation (also when?) Flow at certain gages? satellite data? Non time-variant variables?

# Download data

```{r}
# Required packages to include
# install.packages("librarian")
librarian::shelf(readr, dplyr, tidyr, here, leaflet, sf, ggplot2, ggspatial, prettymapr, sharpshootR, purrr, tibble, ckanr)
```

![[SCHISM boundary conditions](https://cadwrdeltamodeling.github.io/BayDeltaSCHISM/topics/baydelta_boundaries.html)](https://cadwrdeltamodeling.github.io/BayDeltaSCHISM/_images/open_boundaries.png){width="80%"}

```{r}
# SCHISM data

```

## Download USGS data

- Get instantaneous USGS data (for example, discharge sensor data). Start here: ?read_waterdata_continuous. If you only need the latest value, you can use the function: ?read_waterdata_latest_continuous.

- Get daily USGS data (for example, mean daily discharge). Start here: ?read_waterdata_daily. If you only need the latest value, you can use the function: ?read_waterdata_latest_daily.

- Get discrete USGS groundwater data. Start here: ?read_waterdata_field_measurements

- Get discrete water quality data from a cooperative service that integrates publicly available water-quality data from the USGS, EPA, and over 400 state, federal, tribal, and local agencies. Start here: ?readWQPdata

- Get USGS discrete water quality data. Start here: ?read_waterdata_samples

- Get metadata about USGS time series data, including instantaneous and daily data. Start here: ?read_waterdata_ts_meta

- Find Hydro Network-Linked Data Index (NLDI) data. Start here: ?findNLDI

- Get daily data statistics. Start here: ?read_waterdata_stats_por

- Add a USGS WaterData API token to your R environment. See: https://doi-usgs.github.io/dataRetrieval/articles/read_waterdata_functions.html#api-tokens

- Get the latest status on NWIS and other data access updates. See: https://doi-usgs.github.io/dataRetrieval/articles/Status.html https://api.waterdata.usgs.gov/ogcapi/v0/collections/monitoring-locations

```{r}
# https://github.com/DOI-USGS/dataRetrieval

```

## Download data from CDEC
Station notes:
CDEC Stations

ISD - primary
Precip hourly 2006 - present
Wind speed hourly - 2006 - 2015
Wind dir hourly - 2006 - 2015

ANH - fill in
Wind speed 1995 - present
Solar radiation 2008 - present
(Chl-a 2008 - present)
Air temp 2008 -  present

BLP - 
Has some water stuff but not all

DSJ
Water velocity 2003 - present
Discharge 2003 - present
Water temp 2009 - present
EC 2009 - present
Turbidity 2009 - present

SJJ
Water velocity 2006 - present
Discharge 2003 - present
Water temp 2009 - present
EC 2009 - present
Turbidity 2009 - present
DO 2013 - present
Suspended sed 2025 - present

### Pull metadata for CDEC stations
```{r}
#Site investigations for Big Break
ISD_info <- CDEC_StationInfo("ISD") # no water quality data
SJJ_info <- CDEC_StationInfo("SJJ") # Transitioned to USBR from USGS "December 2025"
JER_info <- CDEC_StationInfo("JER") # No comments, is the data on USGS water quality?
BLP_info <- CDEC_StationInfo("BLP") # Blind Point does have WQ data available on DWR Continuous Water Data Library station id = B9502900
ANC_info <- CDEC_StationInfo("ANC") # Becomes ANH in 2018 but link to download data does not work
ANH_info <- CDEC_StationInfo("ANH") # Antioch does not exist?
FRP_info <- CDEC_StationInfo("FRP") # Farrar Park salinity and EC only
BIF_info <- CDEC_StationInfo("BIF") # Precip gage 2008 to present
```
```{r}
#Sensor codes of interest in CDEC 
precipID <- "16"
windID <- "9"
windDirID <- "10"
solarID <- "26" #hourly
airtID <- "4"
watertempID <- "25" #daily
dischargeID <- "20" #event
velocityID <- "21" 
ECID <- "100" #daily or event
turbID <- "221"
DO_ID <- "61"
SSC_ID <- "63"
```

### STATION ID: ISD (Michelle)
```{r}
#STATION ID: ISD Michelle
#Ids <- list("16", "9", "10") 
ISD_PPT <- CDECquery("ISD", "16", "E","2006-11-09", "2026-07-15") 
ISD_WIND <- CDECquery("ISD", "9", "E","2006-11-09", "2026-07-15")
ISD_WINDDIR <- CDECquery("ISD", "10", "E","2006-11-09", "2026-07-15")

plot(ISD_PPT$datetime,ISD_PPT$value)
plot(ISD_WIND$datetime,ISD_WIND$value)
plot(ISD_WINDDIR$datetime,ISD_WINDDIR$value)
```

### STATTION ID: BLP (Michelle)
```{r}
#STATION ID: BLP Michelle
#EC, water temp, DO, pH, turbidity
BLP_EC <- CDECquery("BLP", "100", "D","2010-02-17", "2026-07-15") 
BLP_WaterT <- CDECquery("BLP", "25", "E","2010-02-17", "2026-07-15")
BLP_DO <- CDECquery("BLP", "61", "E","2010-02-17", "2026-07-15")
BLP_pH <- CDECquery("BLP", "62", "E","2010-02-17", "2026-07-15")
BLP_Turb <- CDECquery("BLP", "221", "E","2010-02-17", "2026-07-15")

plot(BLP_EC$datetime,BLP_EC$value)
plot(BLP_WaterT$datetime,BLP_WaterT$value)
plot(BLP_DO$datetime,BLP_DO$value)
plot(BLP_pH$datetime,BLP_pH$value)
plot(BLP_Turb$datetime,BLP_Turb$value)
```

Pull QA QC data from BLP specifically per hidden comment in metatdata [WIP]
```{r}
library(readr)
library(dplyr)

csv_url <- "https://data.ca.gov/datastore/dump/66a144b8-76d4-463e-ba7e-d4a09e715ec7?format=csv"
csv_url2 <- "https://wdlstorageaccount.blob.core.windows.net/wdlcontinuous/B9502900/por/B9502900_WaterTemp_DAYMEAN.csv"
csv_url3 <- "https://wdlstorageaccount.blob.core.windows.net/wdlcontinuous/B9502900/por/B9502900_ECat25C_DAYMEAN.csv"

# Read directly from the CSV dump
df <- read_csv(csv_url)

# Filter for a specific station and daily mean
station_id <- "B9502900"
daily_df <- df %>%
  filter(station_number == station_id, output_interval == "DAYMEAN")

print(head(daily_df))

# Column in daily_df that contains the CSV URLs
csv_url_col <- "download_link"  # <-- change to your actual column name

# --- Function to download and read one CSV ---
read_station_csv <- function(url) {
  message("Downloading: ", url)
  tryCatch({
    read_csv(url, show_col_types = FALSE)
  }, error = function(e) {
    warning("Failed to read ", url, ": ", e$message)
    return(NULL)
  })
}

# --- Download and combine all CSVs ---
all_data <- daily_df %>%
  pull(!!csv_url_col) %>%       # extract the URLs
  map(read_station_csv) #%>%     # download each
  compact() %>%                 # remove NULLs from failed downloads
  bind_rows()                   # stack into one long table

# --- Convert to wide format ---
# Assumes columns: DATE, PARAMETER, VALUE
wide_data <- all_data %>%
  mutate(DATE = as.Date(DATE)) %>%
  pivot_wider(
    id_cols = DATE,
    names_from = PARAMETER,
    values_from = VALUE
  ) %>%
  arrange(DATE)

#View output
head(wide_data)
# --- Export to CSV ---
write_csv(wide_data, "B9502900_DAYMEAN_combined.csv")
```

### STATION ID: SJJ/JER (Michelle)
```{r}
#STATION ID: SJJ/JER Michelle
# Water velocity 2006 - present
# Discharge 2003 - present
# Water temp 2009 - present
# EC 2009 - present
# Turbidity 2009 - present
# DO 2013 - present
# Suspended sed 2025 - present

#EC, water temp, DO, pH, turbidity

SJJ_vel <- CDECquery("SJJ", "21", "E","2006-01-01", "2026-07-15") 
SJJ_Q <- CDECquery("SJJ", "20", "E","2003-01-01", "2026-07-15")
SJJ_WaterT <- CDECquery("SJJ", "25", "E","2009-01-01", "2026-07-15")
SJJ_EC <- CDECquery("SJJ", "100", "E","2009-01-01", "2026-07-15") 
SJJ_Turb <- CDECquery("SJJ", "221", "E","2009-01-01", "2026-07-15")
SJJ_DO <- CDECquery("SJJ", "61", "E","2013-01-01", "2026-07-15")
SJJ_SSC <- CDECquery("SJJ", "63", "E","2025-01-01", "2026-07-15")

plot(SJJ_vel$datetime,SJJ_vel$value)
plot(SJJ_Q$datetime,SJJ_Q$value)
plot(SJJ_WaterT$datetime,SJJ_WaterT$value)
plot(SJJ_EC$datetime,SJJ_EC$value)
plot(SJJ_Turb$datetime,SJJ_Turb$value)
plot(SJJ_DO$datetime,SJJ_DO$value)
plot(SJJ_SSC$datetime,SJJ_SSC$value)

sjj_params <- tibble(
  station = "SJJ",
  sensor = c(21,20,25,100,221,61,63),
  interval = rep("E", 7),
  start = rep("2006-5-22", 7), 
  end = rep("2026-07-15", 7)
)


sjj_all <- sjj_params %>%
  pmap(
    \(station, sensor, interval, start, end) {
      CDECquery(
        id = station,
        sensor = sensor,
        interval = interval,
        start = start,
        end = end
      )
    }
  ) %>%
  list_rbind()
```


### STATION ID: DSJ (Nick)
```{r}
#STATION ID: DSJ Nick

```


### STATION ID: ANH (Nick)
Params:
wind speed (9) - event (hourly, 15 mins) - 2008-10-16 to present
wind direction (10) - event (hourly, 15 mins) - 2008-10-16 to present
solar radiation (26) - event (hourly, 15 mins) - 2008-10-16 to present
air temp (4) - event (hourly, 15 mins) - 2008-10-16 to present
water temp (25) - event (hourly, 15 mins) - 2008-10-16 to present
EC (100) - event (hourly, 15 mins) - 2008-10-16 to present
turbidiy NTU (27) - event (hourly, 15 mins) - 2008-10-16 to present
DO (61) - event (hourly, 15 mins) - 2008-10-16 to present
```{r}
# test one day of wind speed to see how event is handled when it includes hourly and 15 in data ... and now testing more stuff because values returned via this query differ to those on live site
# NB I think event means 15 mins 
anh_windsp_test <- CDECquery("ANH", 9, "E", "2026-07-18", "2026-07-19") # results do not match what is on CDEC site!!!
anh_windsp_test2 <- CDECquery("ANH", 9, "H", "2026-07-18", "2026-07-19")
anh_temp_w <- CDECquery("ANH", 25, "E", "2026-07-18", Sys.Date())
anh_wind_dir <- CDECquery("ANH", 10, "H", "2026-07-18", Sys.Date())

#make event wind speed daily - avg
anh_wind_daily_e <- anh_windsp_test %>%
  mutate(date = as.Date(datetime)) %>%  # extract date
  group_by(date) %>%
  summarize(
    daily_avg = mean(value, na.rm = TRUE),  # avg per day
    .groups = "drop"
  )

plot(anh_wind_daily_e$date,anh_wind_daily_e$daily_avg)

#make hourly wind speed daily - avg
anh_wind_daily_h <- anh_windsp_test2 %>%
  mutate(date = as.Date(datetime)) %>%  # extract date
  group_by(date) %>%
  summarize(
    daily_avg = mean(value, na.rm = TRUE),  # avg per day
    .groups = "drop"
  )

plot(anh_wind_daily_e$daily_avg,anh_wind_daily_h$daily_avg)
anh_wind_daily_e
anh_wind_daily_h

anh_params <- tibble(
  station = "ANH",
  sensor = c(9, 10, 26, 4, 25, 100, 27, 61),
  interval = rep("E", 8),
  start = rep("2008-10-16", 8), 
  end = rep("2026-07-16", 8)
)

anh_all <- anh_params %>%
  pmap(
    \(station, sensor, interval, start, end) {
      CDECquery(
        id = station,
        sensor = sensor,
        interval = interval,
        start = start,
        end = end
      )
    }
  ) %>%
  list_rbind()

```

# Transform data
## Convert cumulative ppt to incremental
```{r}
#convert cumulative precipitation to incremental
ISD_PPT <- ISD_PPT %>% 
  mutate(
    incremental_ppt = c(value[1], diff(value))
  ) 
# remove negative values and anomalously high values greater than 2 inches per hour
#ISD_PPT <- ISD_PPT %>% 
ISD_PPT$incremental_ppt[ISD_PPT$incremental_ppt < 0 ] <- 0
ISD_PPT$incremental_ppt[ISD_PPT$incremental_ppt > 2 ] <- 0

plot(ISD_PPT$datetime,ISD_PPT$incremental_ppt)
```

## Convert hourly incremental ppt to daily
```{r}
#make hourly ppt daily - sum
ISD_PPT_daily <- ISD_PPT %>%
  mutate(date = as.Date(datetime)) %>%  # extract date
  group_by(date) %>%
  summarize(
    daily_sum = sum(incremental_ppt, na.rm = TRUE),  # sum per day
    .groups = "drop"
  )

plot(ISD_PPT_daily$date,ISD_PPT_daily$daily_sum)
```

# Visualize data

## Map CDEC stations
```{r}
#pull station locations
#ANH, ISD, BLP, DSJ, SJJ
#ANH was ANC until 2018
# ISD <- CDEC_StationInfo("ISD")
# BLP <- CDEC_StationInfo("BLP")
# SJJ <- CDEC_StationInfo("SJJ")
# CDEC_StationInfo('ANC')
# CDEC_StationInfo("DSJ")

#grab all Delta stations
library(CDECRetrieve) #had to use the following text to install: devtools::install_github("flowwest/CDECRetrieve", force = TRUE)
Delta_stations_CDEC <- cdec_stations(river_basin = "DELTA") 
Delta_stations_CDEC_sf <-  st_as_sf(Delta_stations_CDEC, coords = c("longitude", "latitude"), crs = 4326)
# Define output shapefile path
out_path <- "data/Delta_CDEC_points.shp"

# Write the sf object to a shapefile
# 'delete_dsn = TRUE' overwrites if file exists
st_write(Delta_stations_CDEC_sf, out_path, delete_dsn = TRUE)
```

```{r}
# gather shapefiles for interactive map of Delta
delta_shapefile_url <- "https://utility.arcgis.com/usrsvcs/servers/3efc635b27344a3da989ca1e7108f5e0/rest/services/Boundaries/i03_LegalDeltaBoundary/MapServer/0/query?outFields=*&where=1%3D1&f=geojson"

# Read the GeoJSON directly from the URL
# geojsonio::geojson_read returns a Spatial object
delta_legal <- st_read(delta_shapefile_url, quiet = TRUE)
# Define projection
delta_legal <- st_transform(delta_legal, crs = 4326)
```

NB some of our stations are missing...
```{r}
# interactive map of Delta with leaflet
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap tiles
  setView(lng = mean(st_coordinates(st_centroid(st_union(delta_legal)))[,1]),
          lat = mean(st_coordinates(st_centroid(st_union(delta_legal)))[,2]),
          zoom = 10) %>%
            #add legal Delta boundary
  addPolygons(
    data = delta_legal,
    color = "blue",
    weight = 2,
    opacity = 0.1,
    fillOpacity = 0.1
    #popup = ~as.character("Legal Delta") # can change to a shapefile column name
  ) %>% 
   #add the CDEC stations
  addCircleMarkers(data = Delta_stations_CDEC_sf,
  color = "red",
  weight = 1.5,
  fillOpacity = 0.4,
  popup = ~paste("Station:", station_id, name)
  )

```



# Merge data together

# Export to format needed for modeling
