---
title: "Download and wrangle hydro datasets"
format: 
  html:
    toc: true
    number-sections: true
    embed-resources: true
authors: Michelle Stern and Nick Rowlands
---

# Brainstorm

The primary environmental factors shaping phytoplankton abundance and community composition are salinity, hydrodynamics, stratification, light, temperature, nutrients, and time. (https://essopenarchive.org/doi/pdf/10.22541/essoar.176495220.03206867/v1) \* Positive: high nutrients, low turbulence, long residence time, adequate light, warm temperatures, low grazing rates, seed population/refugia, low biodiversity. \* Modulating Factors: strong environmental gradients like stratification and salinity, nutrient frms and ratios, diverse habitats/connectivity, selective grazing, predator/prey co-evolution, adequate light, toxin production \* Negative: low nutrients\*, high turbulence, short residence time, low light, high grazing rates, high biodiversity, extreme temperatures warm or cold

Major steps:
1 download raw data
2 QC data, visualize
3 convert to daily
4 combine all parameters together
5 combine with chll data from Emily?
6 combine with toxicity data at Big Break
7 combine with remote sensing indices

# Download data
```{r}
# Required packages to include
# install.packages("librarian")
librarian::shelf(readr, dplyr, tidyr, here, leaflet, sf, ggplot2, ggspatial, prettymapr, sharpshootR, purrr, tibble, ckanr, lubridate)
```

## Download data from CDEC
Station notes:
CDEC Stations

ISD - primary
Precip hourly 2006 - present
Wind speed hourly - 2006 - 2015
Wind dir hourly - 2006 - 2015

ANH - fill in
Wind speed 1995 - present
Solar radiation 2008 - present
(Chl-a 2008 - present)
Air temp 2008 -  present

BLP - 
Has some water stuff but not all

DSJ
Water velocity 2003 - present
Discharge 2003 - present
Water temp 2009 - present
EC 2009 - present
Turbidity 2009 - present

SJJ
Water velocity 2006 - present
Discharge 2003 - present
Water temp 2009 - present
EC 2009 - present
Turbidity 2009 - present
DO 2013 - present
Suspended sed 2025 - present

### Pull metadata for CDEC stations
```{r}
#Site investigations for Big Break
ISD_info <- CDEC_StationInfo("ISD") # no water quality data
SJJ_info <- CDEC_StationInfo("SJJ") # Transitioned to USBR from USGS "December 2025"
JER_info <- CDEC_StationInfo("JER") # No comments, is the data on USGS water quality?
BLP_info <- CDEC_StationInfo("BLP") # Blind Point does have WQ data available on DWR Continuous Water Data Library station id = B9502900
ANC_info <- CDEC_StationInfo("ANC") # Becomes ANH in 2018 but link to download data does not work
ANH_info <- CDEC_StationInfo("ANH") # Antioch does not exist?
FRP_info <- CDEC_StationInfo("FRP") # Farrar Park salinity and EC only
BIF_info <- CDEC_StationInfo("BIF") # Precip gage 2008 to present
#DSJ_info <- CDEC_StationInfo("DSJ")
```
```{r}
#Sensor codes of interest in CDEC 
precipID <- "16"
windID <- "9"
windDirID <- "10"
solarID <- "26" #hourly
airtID <- "4"
watertempID <- "25" #daily
dischargeID <- "20" #event
velocityID <- "21" 
ECID <- "100" #daily or event
turbID <- "221"
DO_ID <- "61"
SSC_ID <- "63"
```

### STATION ID: ISD (Michelle)
```{r}
#STATION ID: ISD Michelle
#Ids <- list("16", "9", "10") 
ISD_PPT <- CDECquery("ISD", "16", "E","2006-11-09", "2026-07-15") 
ISD_WIND <- CDECquery("ISD", "9", "E","2006-11-09", "2026-07-15")
ISD_WINDDIR <- CDECquery("ISD", "10", "E","2006-11-09", "2026-07-15")

plot(ISD_PPT$datetime,ISD_PPT$value)
plot(ISD_WIND$datetime,ISD_WIND$value)
plot(ISD_WINDDIR$datetime,ISD_WINDDIR$value)
```

### STATTION ID: BLP (Michelle)
```{r}
#STATION ID: BLP Michelle
#EC, water temp, DO, pH, turbidity
BLP_EC <- CDECquery("BLP", "100", "D","2010-02-17", "2026-07-15") 
BLP_WaterT <- CDECquery("BLP", "25", "E","2010-02-17", "2026-07-15")
BLP_DO <- CDECquery("BLP", "61", "E","2010-02-17", "2026-07-15")
BLP_pH <- CDECquery("BLP", "62", "E","2010-02-17", "2026-07-15")
BLP_Turb <- CDECquery("BLP", "221", "E","2010-02-17", "2026-07-15")

plot(BLP_EC$datetime,BLP_EC$value)
plot(BLP_WaterT$datetime,BLP_WaterT$value)
plot(BLP_DO$datetime,BLP_DO$value)
plot(BLP_pH$datetime,BLP_pH$value)
plot(BLP_Turb$datetime,BLP_Turb$value)
```

Pull QA QC data from BLP specifically per hidden comment in metatdata [WIP]
```{r}
csv_url <- "https://data.ca.gov/datastore/dump/66a144b8-76d4-463e-ba7e-d4a09e715ec7?format=csv"
csv_url2 <- "https://wdlstorageaccount.blob.core.windows.net/wdlcontinuous/B9502900/por/B9502900_WaterTemp_DAYMEAN.csv"
csv_url3 <- "https://wdlstorageaccount.blob.core.windows.net/wdlcontinuous/B9502900/por/B9502900_ECat25C_DAYMEAN.csv"

# Read directly from the CSV dump
df <- read_csv(csv_url)

# Filter for a specific station and daily mean
station_id <- "B9502900"
daily_df <- df %>%
  filter(station_number == station_id, output_interval == "DAYMEAN")

print(head(daily_df))

# Column in daily_df that contains the CSV URLs
csv_url_col <- "download_link"  # <-- change to your actual column name

# --- Function to download and read one CSV ---
read_station_csv <- function(url) {
  message("Downloading: ", url)
  tryCatch({
    read_csv(url, show_col_types = FALSE)
  }, error = function(e) {
    warning("Failed to read ", url, ": ", e$message)
    return(NULL)
  })
}

# --- Download and combine all CSVs ---
all_data <- daily_df %>%
  pull(!!csv_url_col) %>%       # extract the URLs
  map(read_station_csv) #%>%     # download each
  compact() %>%                 # remove NULLs from failed downloads
  bind_rows()                   # stack into one long table

# --- Convert to wide format ---
# Assumes columns: DATE, PARAMETER, VALUE
wide_data <- all_data %>%
  mutate(DATE = as.Date(DATE)) %>%
  pivot_wider(
    id_cols = DATE,
    names_from = PARAMETER,
    values_from = VALUE
  ) %>%
  arrange(DATE)

#View output
head(wide_data)
# --- Export to CSV ---
write_csv(wide_data, "B9502900_DAYMEAN_combined.csv")
```

### STATION ID: SJJ/JER (Michelle)
```{r}
#STATION ID: SJJ/JER Michelle
# Water velocity 2006 - present
# Discharge 2003 - present
# Water temp 2009 - present
# EC 2009 - present
# Turbidity 2009 - present
# DO 2013 - present
# Suspended sed 2025 - present

#EC, water temp, DO, pH, turbidity

SJJ_vel <- CDECquery("SJJ", "21", "E","2006-01-01", "2026-07-15") 
SJJ_Q <- CDECquery("SJJ", "20", "E","2003-01-01", "2026-07-15")
SJJ_WaterT <- CDECquery("SJJ", "25", "E","2009-01-01", "2026-07-15")
SJJ_EC <- CDECquery("SJJ", "100", "E","2009-01-01", "2026-07-15") 
SJJ_Turb <- CDECquery("SJJ", "221", "E","2009-01-01", "2026-07-15")
SJJ_DO <- CDECquery("SJJ", "61", "E","2013-01-01", "2026-07-15")
SJJ_SSC <- CDECquery("SJJ", "63", "E","2025-01-01", "2026-07-15")

plot(SJJ_vel$datetime,SJJ_vel$value)
plot(SJJ_Q$datetime,SJJ_Q$value)
plot(SJJ_WaterT$datetime,SJJ_WaterT$value)
plot(SJJ_EC$datetime,SJJ_EC$value)
plot(SJJ_Turb$datetime,SJJ_Turb$value)
plot(SJJ_DO$datetime,SJJ_DO$value)
plot(SJJ_SSC$datetime,SJJ_SSC$value)

sjj_params <- tibble(
  station = "SJJ",
  sensor = c(21,20,25,100,221,61,63),
  interval = rep("E", 7),
  start = rep("2006-5-22", 7), 
  end = rep("2026-07-15", 7)
)


sjj_all <- sjj_params %>%
  pmap(
    \(station, sensor, interval, start, end) {
      CDECquery(
        id = station,
        sensor = sensor,
        interval = interval,
        start = start,
        end = end
      )
    }
  ) %>%
  list_rbind()
```


### STATION ID: DSJ (Nick)
```{r}
#STATION ID: DSJ Nick

```


### STATION ID: ANH (Nick)
Params:
wind speed (9) - event (hourly, 15 mins) - 2008-10-16 to present
wind direction (10) - event (hourly, 15 mins) - 2008-10-16 to present
solar radiation (26) - event (hourly, 15 mins) - 2008-10-16 to present
air temp (4) - event (hourly, 15 mins) - 2008-10-16 to present
water temp (25) - event (hourly, 15 mins) - 2008-10-16 to present
EC (100) - event (hourly, 15 mins) - 2008-10-16 to present
turbidiy NTU (27) - event (hourly, 15 mins) - 2008-10-16 to present
DO (61) - event (hourly, 15 mins) - 2008-10-16 to present
```{r}
# test one day of wind speed to see how event is handled when it includes hourly and 15 in data ... and now testing more stuff because values returned via this query differ to those on live site
# NB I think event means 15 mins 
anh_windsp_test <- CDECquery("ANH", 9, "E", "2026-07-18", "2026-07-19") # results do not match what is on CDEC site!!!
anh_windsp_test2 <- CDECquery("ANH", 9, "H", "2026-07-18", "2026-07-19")
anh_temp_w <- CDECquery("ANH", 25, "E", "2026-07-18", Sys.Date())
anh_wind_dir <- CDECquery("ANH", 10, "H", "2026-07-18", Sys.Date())

#make event wind speed daily - avg
anh_wind_daily_e <- anh_windsp_test %>%
  mutate(date = as.Date(datetime)) %>%  # extract date
  group_by(date) %>%
  summarize(
    daily_avg = mean(value, na.rm = TRUE),  # avg per day
    .groups = "drop"
  )

plot(anh_wind_daily_e$date,anh_wind_daily_e$daily_avg)

#make hourly wind speed daily - avg
anh_wind_daily_h <- anh_windsp_test2 %>%
  mutate(date = as.Date(datetime)) %>%  # extract date
  group_by(date) %>%
  summarize(
    daily_avg = mean(value, na.rm = TRUE),  # avg per day
    .groups = "drop"
  )

plot(anh_wind_daily_e$daily_avg,anh_wind_daily_h$daily_avg)
anh_wind_daily_e
anh_wind_daily_h

anh_params <- tibble(
  station = "ANH",
  sensor = c(9, 10, 26, 4, 25, 100, 27, 61),
  interval = rep("E", 8),
  start = rep("2008-10-16", 8), 
  end = rep("2026-07-16", 8)
)

anh_all <- anh_params %>%
  pmap(
    \(station, sensor, interval, start, end) {
      CDECquery(
        id = station,
        sensor = sensor,
        interval = interval,
        start = start,
        end = end
      )
    }
  ) %>%
  list_rbind()

```

#### Pull ANH data from QC'd dataset shared by Andrew Tran @ DWR
```{r}
# Package ID: edi.1177.8 Cataloging System:https://pasta.edirepository.org.
# Data set title: Sacramento-San Joaquin Bay-Delta Continuous (15 Minute) water quality monitoring data collected by the Continuous Environmental Monitoring Program, DWR, 2005- 2025..
# Data set creator:  Michelle Nelson - Department of Water Resources 
# Data set creator:  Andrew Tran - Department of Water Resources 
#  Metadata Provider:  Michelle Nelson - Department of Water Resources 
#  Metadata Provider:  Andrew Tran - Department of Water Resources 
#  Contact:  Michelle Nelson - Environmental Scientist Department of Water Resources  - Michelle.Nelson@water.ca.gov
#  Contact:  Andrew Tran - Environmental Scientist Department of Water Resources  - Andrew.Tran@water.ca.gov 
# You should replace 'PUT-LOCAL-PATH-TO-DATA-FILE-HERE'  (below) with the appropriate path.   
#   to your data file (e.g., c:\mydata\datafile.txt). 
infile1 <- trimws("https://pasta.lternet.edu/package/data/eml/edi/1177/8/a716ea3db39bb1ba35213ccc4169bebf")
infile1 <-sub("^https","http",infile1) 
 dataTable1 <-read.csv(infile1, 
            ,sep=","  
                ,quot='"' 
        , col.names=c(
                    "station",     
                    "description",     
                    "station_access_type",     
                    "latitude",     
                    "longitude",     
                    "hydrologic_area",     
                    "basin",     
                    "county",     
                    "habitat_type",     
                    "start_date",     
                    "end_date",     
                    "EXO2_start_date",     
                    "EXO2_start_time",     
                    "notes"    ), check.names=TRUE)

               

 
tmp_var <- character()
tmp_label <- character()
                       tmp_var <- c(tmp_var , "station") 
                       tmp_label <- c(tmp_label , "Station Acronym- ")
                       tmp_var <- c(tmp_var , "description") 
                       tmp_label <- c(tmp_label , "StationName- ")
                       tmp_var <- c(tmp_var , "station_access_type") 
                       tmp_label <- c(tmp_label , "Station Access Type- ")
                       tmp_var <- c(tmp_var , "latitude") 
                       tmp_label <- c(tmp_label , "Latitude- degree")
                       tmp_var <- c(tmp_var , "longitude") 
                       tmp_label <- c(tmp_label , "Longitude- degree")
                       tmp_var <- c(tmp_var , "hydrologic_area") 
                       tmp_label <- c(tmp_label , "HydrologicArea- ")
                       tmp_var <- c(tmp_var , "basin") 
                       tmp_label <- c(tmp_label , "Basin- ")
                       tmp_var <- c(tmp_var , "county") 
                       tmp_label <- c(tmp_label , "County- ")
                       tmp_var <- c(tmp_var , "habitat_type") 
                       tmp_label <- c(tmp_label , "HabitatType- ")
                       tmp_var <- c(tmp_var , "start_date") 
                       tmp_label <- c(tmp_label , "Dataset Begin Date- ")
                       tmp_var <- c(tmp_var , "end_date") 
                       tmp_label <- c(tmp_label , "Dataset End Date- ")
                       tmp_var <- c(tmp_var , "EXO2_start_date") 
                       tmp_label <- c(tmp_label , "Start date of YSI EXO2 date- ")
                       tmp_var <- c(tmp_var , "EXO2_start_time") 
                       tmp_label <- c(tmp_label , "YSI EXO2 start time- ")
                       tmp_var <- c(tmp_var , "notes") 
                       tmp_label <- c(tmp_label , "Notes- ")
labelFrame1<-data.frame(variable=tmp_var, label=tmp_label)
rm(tmp_var, tmp_label)        
 

tmp_var <- character()
tmp_code <- character()                
tmp_label <- character() 
tmp_var <- c(tmp_var,"station_access_type")   
tmp_code <- c(tmp_code,"Land")   
tmp_label <- c(tmp_label,"Land sites are accessed by vehicle and are located on the banks of river channels.")  
tmp_var <- c(tmp_var,"station_access_type")   
tmp_code <- c(tmp_code,"Pile")   
tmp_label <- c(tmp_label,"Pile sites are accessed by boats and are in the middle of river channels.")  
tmp_var <- c(tmp_var,"hydrologic_area")   
tmp_code <- c(tmp_code,"Sacramento River")   
tmp_label <- c(tmp_label,"Sacramento River")  
tmp_var <- c(tmp_var,"hydrologic_area")   
tmp_code <- c(tmp_code,"San Francisco Bay")   
tmp_label <- c(tmp_label,"San Francisco Bay")  
tmp_var <- c(tmp_var,"hydrologic_area")   
tmp_code <- c(tmp_code,"San Joaquin River")   
tmp_label <- c(tmp_label,"San Joaquin River")  
tmp_var <- c(tmp_var,"basin")   
tmp_code <- c(tmp_code,"Delta")   
tmp_label <- c(tmp_label,"within the legal limits of the Sacramento- San Joaquin Delta")  
tmp_var <- c(tmp_var,"basin")   
tmp_code <- c(tmp_code,"East Bay")   
tmp_label <- c(tmp_label,"Eastern bay on the Western side of the Sacramento- San Joaquin Delta legal boundary")  
tmp_var <- c(tmp_var,"basin")   
tmp_code <- c(tmp_code,"Sacramento R")   
tmp_label <- c(tmp_label,"Sacramento River")  
tmp_var <- c(tmp_var,"basin")   
tmp_code <- c(tmp_code,"San Joaquin R")   
tmp_label <- c(tmp_label,"San Joaquin River")  
tmp_var <- c(tmp_var,"county")   
tmp_code <- c(tmp_code,"Contra Costa")   
tmp_label <- c(tmp_label,"Contra Costa County")  
tmp_var <- c(tmp_var,"county")   
tmp_code <- c(tmp_code,"Sacramento")   
tmp_label <- c(tmp_label,"Sacramento County")  
tmp_var <- c(tmp_var,"county")   
tmp_code <- c(tmp_code,"San Joaquin")   
tmp_label <- c(tmp_label,"San Joaquin County")  
tmp_var <- c(tmp_var,"county")   
tmp_code <- c(tmp_code,"Solano")   
tmp_label <- c(tmp_label,"Solano County")  
tmp_var <- c(tmp_var,"habitat_type")   
tmp_code <- c(tmp_code,"Tidal Embayment (brackish)")   
tmp_label <- c(tmp_label,"A recess in a waterway or coastline forming a bay with brackish water.")  
tmp_var <- c(tmp_var,"habitat_type")   
tmp_code <- c(tmp_code,"Tidal Flooded Island (freshwater)")   
tmp_label <- c(tmp_label,"A man-made flooded island that contains freshwater.")  
tmp_var <- c(tmp_var,"habitat_type")   
tmp_code <- c(tmp_code,"Tidal River Channel (brackish)")   
tmp_label <- c(tmp_label,"A natural brackish river (higher salinity than freshwater, resulting from a mixture of estuarine river and seawater) whose flow and level are influenced by tides.")  
tmp_var <- c(tmp_var,"habitat_type")   
tmp_code <- c(tmp_code,"Tidal River Channel (freshwater)")   
tmp_label <- c(tmp_label,"A natural river, mainly freshwater, whose flow and level are influenced by tides.")  
tmp_var <- c(tmp_var,"habitat_type")   
tmp_code <- c(tmp_code,"Tidal River Channel (Freshwater)")   
tmp_label <- c(tmp_label,"Main river channel that is freshwater")  
codeLabelFrame1 <- data.frame(variable=tmp_var, code=tmp_code, label=tmp_label) 
rm(tmp_var, tmp_label, tmp_code) 
                
# You should replace 'PUT-LOCAL-PATH-TO-DATA-FILE-HERE'  (below) with the appropriate path.   
#   to your data file (e.g., c:\mydata\datafile.txt). 
infile3 <- trimws("https://pasta.lternet.edu/package/data/eml/edi/1177/8/e7c35c12cd774b2146e596f2303872f7")
infile3 <-sub("^https","http",infile3) 
 ANH_QC15 <-read.csv(infile3, 
            ,sep=","  
                ,quot='"' 
        , col.names=c(
                    "station",     
                    "date",     
                    "time",     
                    "dissolvedoxygen",     
                    "fluorescence",     
                    "ph",     
                    "spc",     
                    "turbidity",     
                    "watertemperature"    ), check.names=TRUE)

              
tmp_var <- character()
tmp_label <- character()
                       tmp_var <- c(tmp_var , "station") 
                       tmp_label <- c(tmp_label , "station- ")
                       tmp_var <- c(tmp_var , "date") 
                       tmp_label <- c(tmp_label , "date- ")
                       tmp_var <- c(tmp_var , "time") 
                       tmp_label <- c(tmp_label , "time- ")
                       tmp_var <- c(tmp_var , "dissolvedoxygen") 
                       tmp_label <- c(tmp_label , "dissolvedoxygen- milligramPerLiter")
                       tmp_var <- c(tmp_var , "fluorescence") 
                       tmp_label <- c(tmp_label , "fluorescence- microgramPerLiter")
                       tmp_var <- c(tmp_var , "ph") 
                       tmp_label <- c(tmp_label , "ph- pH")
                       tmp_var <- c(tmp_var , "spc") 
                       tmp_label <- c(tmp_label , "spc- microsiemens per centimeter")
                       tmp_var <- c(tmp_var , "turbidity") 
                       tmp_label <- c(tmp_label , "turbidity- FNU")
                       tmp_var <- c(tmp_var , "watertemperature") 
                       tmp_label <- c(tmp_label , "watertemperature- celsius")
labelFrame3<-data.frame(variable=tmp_var, label=tmp_label)
rm(tmp_var, tmp_label)        

tmp_var <- character()
tmp_code <- character()                
tmp_label <- character() 
tmp_var <- c(tmp_var,"station")   
tmp_code <- c(tmp_code,"ANH")   
tmp_label <- c(tmp_label,"Station Abbreviation")  
codeLabelFrame3 <- data.frame(variable=tmp_var, code=tmp_code, label=tmp_label) 
rm(tmp_var, tmp_label, tmp_code) 
                
 # HERE IS A LIST OF VARIABLES TO WHICH CODES HAD BEEN ASSIGNED               
codeLabelFrame3           
                
attach(ANH_QC15)               
# The analyses below are basic descriptions of the variables. After testing, they should be replaced.                 
        
                 
summary(as.factor(station))  
summary(as.numeric(dissolvedoxygen))  
summary(as.numeric(fluorescence))  
summary(as.numeric(ph))  
summary(as.numeric(spc))  
summary(as.numeric(turbidity))  
summary(as.numeric(watertemperature))  
```

# QC raw CDEC data and transform to daily
```{r}
# Function to compute descriptive statistics for a time series
ts_summary <- function(x) {
  # Validate input
  if (!is.numeric(x)) {
    stop("Input must be numeric or a time series object.")
  }
  
  # Remove NA values
  x <- na.omit(as.numeric(x))
  
  if (length(x) < 2) {
    stop("Time series must have at least 2 non-missing values.")
  }
  
  # Compute statistics
  stats <- data.frame(
    Count = length(x),
    Mean = mean(x),
    Median = median(x),
    Min = min(x),
    Max = max(x),
    Range = diff(range(x)),
    SD = sd(x),
    Variance = var(x),
    Q05 = quantile(x, 0.05),
    Q95 = quantile(x, 0.95),
    Lag1_ACF = acf(x, plot = FALSE)$acf[2]
  )
  
  return(stats)
}
```

## velocity
```{r}
SJJ_vel <- CDECquery("SJJ", "21", "E","2006-01-01", "2026-07-15") 
# summary statistics and removal of outliers
# run summary stats
ts_summary(SJJ_vel$value)

threshold_pos_vel <- 4
threshold_neg_vel <- -6

# Remove values outside thresholds (set to NA)
SJJ_vel <- SJJ_vel %>%
  mutate(value = ifelse(value > threshold_pos_vel | value < threshold_neg_vel, NA, value))

ts_summary(SJJ_vel$value)
SJJavg <- mean(SJJ_vel$value, na.rm = TRUE)
SJJstd <- sd(SJJ_vel$value, na.rm = TRUE)
plot(SJJ_vel$datetime,SJJ_vel$value)

#testing 
threshold_pos_vel2 <- (SJJavg + 2 * SJJstd)
threshold_neg_vel2 <- (SJJavg - 2 * SJJstd)
SJJ_vel2 <- SJJ_vel %>%
  mutate(value = ifelse(value > threshold_pos_vel | value < threshold_neg_vel, NA, value))
plot(SJJ_vel2$datetime,SJJ_vel2$value)
ts_summary(SJJ_vel2$value)

# Step 1: Aggregate 15-min data to daily mean
SJJ_vel_daily <- SJJ_vel %>%
  mutate(datetime = as.Date(datetime)) %>%
  group_by(datetime) %>%
  summarise(daily_mean = mean(value, na.rm = TRUE), .groups = "drop")

# Step 2: Add year and day-of-year (DOY)
SJJ_vel_daily <- SJJ_vel_daily %>%
  mutate(
    year = year(datetime),
    doy  = yday(datetime)
  )

# Step 3: Calculate standard deviation of daily means for each DOY across years
doy_stats <- SJJ_vel_daily %>%
  group_by(doy) %>%
  summarise(
    mean_daily_mean = mean(daily_mean, na.rm = TRUE),
    sd_daily_mean = sd(daily_mean, na.rm = TRUE),
    n_years = n(),  # optional: number of daily values used
    .groups = "drop"
  )

# View results
head(doy_stats, 10)
plot(SJJ_vel_daily$datetime,SJJ_vel_daily$daily_mean)
plot(doy_stats$doy,doy_stats$mean_daily_mean)
plot(doy_stats$doy,doy_stats$sd_daily_mean)
```

## flow
```{r}
SJJ_Q <- CDECquery("SJJ", "20", "E","2003-01-01", "2026-07-15")
# summary statistics and removal of outliers
plot(SJJ_Q$value)

# run summary stats
ts_summary(SJJ_Q$value)

threshold_pos_Q <- 900000 #1480485
threshold_neg_Q <- -900000 #-1407986

# Remove values outside thresholds (set to NA)
SJJ_Q <- SJJ_Q %>%
  mutate(value = ifelse(value > threshold_pos_Q | value < threshold_neg_Q, NA, value))

ts_summary(SJJ_Q$value)
plot(SJJ_Q$datetime,SJJ_Q$value)

# Step 1: Aggregate 15-min data to daily mean
SJJ_Q_daily <- SJJ_Q %>%
  mutate(datetime = as.Date(datetime)) %>%
  group_by(datetime) %>%
  summarise(daily_mean = mean(value, na.rm = TRUE), .groups = "drop")

# Step 2: Add year and day-of-year (DOY)
SJJ_Q_daily <- SJJ_Q_daily %>%
  mutate(
    year = year(datetime),
    doy  = yday(datetime)
  )

# Step 3: Calculate standard deviation of daily means for each DOY across years
doy_stats <- SJJ_Q_daily %>%
  group_by(doy) %>%
  summarise(
    mean_daily_mean = mean(daily_mean, na.rm = TRUE),
    sd_daily_mean = sd(daily_mean, na.rm = TRUE),
    n_years = n(),  # optional: number of daily values used
    .groups = "drop"
  )

# View results
head(doy_stats, 10)
plot(SJJ_Q_daily$datetime,SJJ_Q_daily$daily_mean)
plot(doy_stats$doy,doy_stats$mean_daily_mean)
plot(doy_stats$doy,doy_stats$sd_daily_mean)
```

## water temp
```{r}
SJJ_WaterT <- CDECquery("SJJ", "25", "E","2009-01-01", "2026-07-15")
# summary statistics and removal of outliers
plot(SJJ_WaterT$value)

# run summary stats
ts_summary(SJJ_WaterT$value)

threshold_pos_WaterT <- 100 #163.8
threshold_neg_WaterT <- 40 #-17966.2

# Remove values outside thresholds (set to NA)
SJJ_WaterT <- SJJ_WaterT %>%
  mutate(value = ifelse(value > threshold_pos_WaterT | value < threshold_neg_WaterT, NA, value))

ts_summary(SJJ_WaterT$value)
plot(SJJ_WaterT$datetime,SJJ_WaterT$value)

# Step 1: Aggregate 15-min data to daily mean
SJJ_WaterT_daily <- SJJ_WaterT %>%
  mutate(datetime = as.Date(datetime)) %>%
  group_by(datetime) %>%
  summarise(daily_mean = mean(value, na.rm = TRUE), .groups = "drop")

# Step 2: Add year and day-of-year (DOY)
SJJ_WaterT_daily <- SJJ_WaterT_daily %>%
  mutate(
    year = year(datetime),
    doy  = yday(datetime)
  )

# Step 3: Calculate standard deviation of daily means for each DOY across years
doy_stats <- SJJ_WaterT_daily %>%
  group_by(doy) %>%
  summarise(
    mean_daily_mean = mean(daily_mean, na.rm = TRUE),
    sd_daily_mean = sd(daily_mean, na.rm = TRUE),
    n_years = n(),  # optional: number of daily values used
    .groups = "drop"
  )

# View results
head(doy_stats, 10)
plot(SJJ_WaterT_daily$datetime,SJJ_WaterT_daily$daily_mean)
plot(doy_stats$doy,doy_stats$mean_daily_mean)
plot(doy_stats$doy,doy_stats$sd_daily_mean)

```

## EC
```{r}
SJJ_EC <- CDECquery("SJJ", "100", "E","2009-01-01", "2026-07-15")
# summary statistics and removal of outliers
plot(SJJ_EC$value)

# run summary stats
ts_summary(SJJ_EC$value)

threshold_pos_EC <- 10000 
threshold_neg_EC <- -10 

# Remove values outside thresholds (set to NA)
SJJ_EC <- SJJ_EC %>%
  mutate(value = ifelse(value > threshold_pos_EC | value < threshold_neg_EC, NA, value))

ts_summary(SJJ_EC$value)
plot(SJJ_EC$datetime,SJJ_EC$value)

# Step 1: Aggregate 15-min data to daily mean
SJJ_EC_daily <- SJJ_EC %>%
  mutate(datetime = as.Date(datetime)) %>%
  group_by(datetime) %>%
  summarise(daily_mean = mean(value, na.rm = TRUE), .groups = "drop")

# Step 2: Add year and day-of-year (DOY)
SJJ_EC_daily <- SJJ_EC_daily %>%
  mutate(
    year = year(datetime),
    doy  = yday(datetime)
  )

# Step 3: Calculate standard deviation of daily means for each DOY across years
doy_stats <- SJJ_EC_daily %>%
  group_by(doy) %>%
  summarise(
    mean_daily_mean = mean(daily_mean, na.rm = TRUE),
    sd_daily_mean = sd(daily_mean, na.rm = TRUE),
    n_years = n(),  # optional: number of daily values used
    .groups = "drop"
  )

# View results
head(doy_stats, 10)
plot(SJJ_EC_daily$datetime,SJJ_EC_daily$daily_mean)
plot(doy_stats$doy,doy_stats$mean_daily_mean)
plot(doy_stats$doy,doy_stats$sd_daily_mean)

```

```{r}
# summary statistics and removal of outliers
plot(BLP_EC$value)

# run summary stats
ts_summary(BLP_EC$value)

threshold_pos_EC <- 10000 
threshold_neg_EC <- 0.1 

# Remove values outside thresholds (set to NA)
BLP_EC <- BLP_EC %>%
  mutate(value = ifelse(value > threshold_pos_EC | value < threshold_neg_EC, NA, value))

ts_summary(BLP_EC$value)
plot(BLP_EC$datetime,BLP_EC$value)

# Step 1: Aggregate 15-min data to daily mean
BLP_EC_daily <- BLP_EC %>%
  mutate(datetime = as.Date(datetime)) %>%
  group_by(datetime) %>%
  summarise(daily_mean = mean(value, na.rm = TRUE), .groups = "drop")

# Step 2: Add year and day-of-year (DOY)
BLP_EC_daily <- BLP_EC_daily %>%
  mutate(
    year = year(datetime),
    doy  = yday(datetime)
  )

# Step 3: Calculate standard deviation of daily means for each DOY across years
BLP_EC_doy_stats <- BLP_EC_daily %>%
  group_by(doy) %>%
  summarise(
    mean_daily_mean = mean(daily_mean, na.rm = TRUE),
    sd_daily_mean = sd(daily_mean, na.rm = TRUE),
    n_years = n(),  # optional: number of daily values used
    .groups = "drop"
  )

# View results
head(BLP_EC_doy_stats, 10)
plot(BLP_EC_daily$datetime,BLP_EC_daily$daily_mean)
plot(BLP_EC_doy_stats$doy,BLP_EC_doy_stats$mean_daily_mean)
plot(BLP_EC_doy_stats$doy,BLP_EC_doy_stats$sd_daily_mean)

All_EC_daily <- full_join(BLP_EC_daily,SJJ_EC_daily, by = "datetime") %>% 
  arrange(datetime) 

names(All_EC_daily)[names(All_EC_daily) == "daily_mean.x"] <- "BLP_EC" 
names(All_EC_daily)[names(All_EC_daily) == "daily_mean.y"] <- "SJJ_EC"

print(All_EC_daily)

# Convert from wide to long format for ggplot
All_EC_daily_long <- All_EC_daily %>%
  pivot_longer(cols = c(BLP_EC, SJJ_EC),
               names_to = "series",
               values_to = "value")

# Plot both series on the same graph
ggplot(All_EC_daily_long, aes(x = datetime, y = value, color = series)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(
    x = "Date",
    y = "EC",
    color = "Site"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    legend.position = "top"
  )
```



## turbidity
```{r}
SJJ_Turb <- CDECquery("SJJ", "221", "E","2009-01-01", "2026-07-15")
# summary statistics and removal of outliers
plot(SJJ_Turb$value)

# run summary stats
ts_summary(SJJ_Turb$value)

threshold_pos_Turb <- 10000 
threshold_neg_Turb <- -10 

# Remove values outside thresholds (set to NA)
SJJ_Turb <- SJJ_Turb %>%
  mutate(value = ifelse(value > threshold_pos_Turb | value < threshold_neg_Turb, NA, value))

ts_summary(SJJ_Turb$value)
plot(SJJ_Turb$datetime,SJJ_Turb$value)

# Step 1: Aggregate 15-min data to daily mean
SJJ_Turb_daily <- SJJ_Turb %>%
  mutate(datetime = as.Date(datetime)) %>%
  group_by(datetime) %>%
  summarise(daily_mean = mean(value, na.rm = TRUE), .groups = "drop")

# Step 2: Add year and day-of-year (DOY)
SJJ_Turb_daily <- SJJ_Turb_daily %>%
  mutate(
    year = year(datetime),
    doy  = yday(datetime)
  )

# Step 3: Calculate standard deviation of daily means for each DOY across years
doy_stats <- SJJ_Turb_daily %>%
  group_by(doy) %>%
  summarise(
    mean_daily_mean = mean(daily_mean, na.rm = TRUE),
    sd_daily_mean = sd(daily_mean, na.rm = TRUE),
    n_years = n(),  # optional: number of daily values used
    .groups = "drop"
  )

# View results
head(doy_stats, 10)
plot(SJJ_Turb_daily$datetime,SJJ_Turb_daily$daily_mean)
plot(doy_stats$doy,doy_stats$mean_daily_mean)
plot(doy_stats$doy,doy_stats$sd_daily_mean)

```

## DO
```{r}
SJJ_DO <- CDECquery("SJJ", "61", "E","2013-01-01", "2026-07-15")
# summary statistics and removal of outliers
plot(SJJ_DO$value)

# run summary stats
ts_summary(SJJ_DO$value)

threshold_pos_DO <- 15 
threshold_neg_DO <- 2 

# Remove values outside thresholds (set to NA)
SJJ_DO <- SJJ_DO %>%
  mutate(value = ifelse(value > threshold_pos_DO | value < threshold_neg_DO, NA, value))

ts_summary(SJJ_DO$value)
plot(SJJ_DO$datetime,SJJ_DO$value)

# Step 1: Aggregate 15-min data to daily mean
SJJ_DO_daily <- SJJ_DO %>%
  mutate(datetime = as.Date(datetime)) %>%
  group_by(datetime) %>%
  summarise(daily_mean = mean(value, na.rm = TRUE), .groups = "drop")

# Step 2: Add year and day-of-year (DOY)
SJJ_DO_daily <- SJJ_DO_daily %>%
  mutate(
    year = year(datetime),
    doy  = yday(datetime)
  )

# Step 3: Calculate standard deviation of daily means for each DOY across years
doy_stats <- SJJ_DO_daily %>%
  group_by(doy) %>%
  summarise(
    mean_daily_mean = mean(daily_mean, na.rm = TRUE),
    sd_daily_mean = sd(daily_mean, na.rm = TRUE),
    n_years = n(),  # optional: number of daily values used
    .groups = "drop"
  )

# View results
head(doy_stats, 10)
plot(SJJ_DO_daily$datetime,SJJ_DO_daily$daily_mean)
plot(doy_stats$doy,doy_stats$mean_daily_mean)
plot(doy_stats$doy,doy_stats$sd_daily_mean)
```

## SSC
```{r}
SJJ_SSC <- CDECquery("SJJ", "63", "E","2025-01-01", "2026-07-15")
# summary statistics and removal of outliers
plot(SJJ_SSC$value)

# run summary stats
ts_summary(SJJ_SSC$value)

threshold_pos_SSC <- 50 
threshold_neg_SSC <- -10 

# Remove values outside thresholds (set to NA)
SJJ_SSC <- SJJ_SSC %>%
  mutate(value = ifelse(value > threshold_pos_SSC | value < threshold_neg_SSC, NA, value))

ts_summary(SJJ_SSC$value)
plot(SJJ_SSC$datetime,SJJ_SSC$value)

# Step 1: Aggregate 15-min data to daily mean
SJJ_SSC_daily <- SJJ_SSC %>%
  mutate(datetime = as.Date(datetime)) %>%
  group_by(datetime) %>%
  summarise(daily_mean = mean(value, na.rm = TRUE), .groups = "drop")

# Step 2: Add year and day-of-year (DOY)
SJJ_SSC_daily <- SJJ_SSC_daily %>%
  mutate(
    year = year(datetime),
    doy  = yday(datetime)
  )

# Step 3: Calculate standard deviation of daily means for each DOY across years
doy_stats <- SJJ_SSC_daily %>%
  group_by(doy) %>%
  summarise(
    mean_daily_mean = mean(daily_mean, na.rm = TRUE),
    sd_daily_mean = sd(daily_mean, na.rm = TRUE),
    n_years = n(),  # optional: number of daily values used
    .groups = "drop"
  )

# View results
head(doy_stats, 10)
plot(SJJ_SSC_daily$datetime,SJJ_SSC_daily$daily_mean)
plot(doy_stats$doy,doy_stats$mean_daily_mean)
# plot(doy_stats$doy,doy_stats$sd_daily_mean)

```

## wind
```{r}
ISD_WIND <- CDECquery("ISD", "16", "E","2006-11-09", "2026-07-15")
# summary statistics and removal of outliers
plot(ISD_WIND$value)

# run summary stats
ts_summary(ISD_WIND$value)

# threshold_pos_SSC <- 50 
# threshold_neg_SSC <- -10 

# # Remove values outside thresholds (set to NA)
# ISD_WIND <- ISD_WIND %>%
#   mutate(value = ifelse(value > threshold_pos_SSC | value < threshold_neg_SSC, NA, value))

ts_summary(ISD_WIND$value)
plot(ISD_WIND$datetime,ISD_WIND$value)

# Step 1: Aggregate 15-min data to daily mean
ISD_WIND_daily <- ISD_WIND %>%
  mutate(datetime = as.Date(datetime)) %>%
  group_by(datetime) %>%
  summarise(daily_mean = mean(value, na.rm = TRUE), .groups = "drop")

# Step 2: Add year and day-of-year (DOY)
ISD_WIND_daily <- ISD_WIND_daily %>%
  mutate(
    year = year(datetime),
    doy  = yday(datetime)
  )

# Step 3: Calculate standard deviation of daily means for each DOY across years
doy_stats <- ISD_WIND_daily %>%
  group_by(doy) %>%
  summarise(
    mean_daily_mean = mean(daily_mean, na.rm = TRUE),
    sd_daily_mean = sd(daily_mean, na.rm = TRUE),
    n_years = n(),  # optional: number of daily values used
    .groups = "drop"
  )

# View results
head(doy_stats, 10)
plot(ISD_WIND_daily$datetime,ISD_WIND_daily$daily_mean)
plot(doy_stats$doy,doy_stats$mean_daily_mean)
plot(doy_stats$doy,doy_stats$sd_daily_mean)
```

## Wind Direction
```{r}
ISD_WINDDIR <- CDECquery("ISD", "16", "E","2006-11-09", "2026-07-15")
# summary statistics and removal of outliers
plot(ISD_WINDDIR$value)

# run summary stats
ts_summary(ISD_WINDDIR$value)

# threshold_pos_WINDDIR <- 50 
# threshold_neg_WINDDIR <- -10 

# # Remove values outside thresholds (set to NA)
# ISD_WINDDIR <- ISD_WINDDIR %>%
#   mutate(value = ifelse(value > threshold_pos_WINDDIR | value < threshold_neg_WINDDIR, NA, value))

ts_summary(ISD_WINDDIR$value)
plot(ISD_WINDDIR$datetime,ISD_WINDDIR$value)

# Step 1: Aggregate 15-min data to daily mean
ISD_WINDDIR_daily <- ISD_WINDDIR %>%
  mutate(datetime = as.Date(datetime)) %>%
  group_by(datetime) %>%
  summarise(daily_mean = mean(value, na.rm = TRUE), .groups = "drop")

# Step 2: Add year and day-of-year (DOY)
ISD_WINDDIR_daily <- ISD_WINDDIR_daily %>%
  mutate(
    year = year(datetime),
    doy  = yday(datetime)
  )

# Step 3: Calculate standard deviation of daily means for each DOY across years
ISD_WINDDIR_doy_stats <- ISD_WINDDIR_daily %>%
  group_by(doy) %>%
  summarise(
    mean_daily_mean = mean(daily_mean, na.rm = TRUE),
    sd_daily_mean = sd(daily_mean, na.rm = TRUE),
    n_years = n(),  # optional: number of daily values used
    .groups = "drop"
  )

# View results
head(ISD_WINDDIR_doy_stats, 10)
plot(ISD_WINDDIR_daily$datetime,ISD_WINDDIR_daily$daily_mean)
plot(ISD_WINDDIR_doy_stats$doy,ISD_WINDDIR_doy_stats$mean_daily_mean)
plot(ISD_WINDDIR_doy_stats$doy,ISD_WINDDIR_doy_stats$sd_daily_mean)
```


## Convert cumulative ppt to incremental
```{r}
#convert cumulative precipitation to incremental
ISD_PPT <- ISD_PPT %>% 
  mutate(
    incremental_ppt = c(value[1], diff(value))
  ) 
# remove negative values and anomalously high values greater than 2 inches per hour
#ISD_PPT <- ISD_PPT %>% 
ISD_PPT$incremental_ppt[ISD_PPT$incremental_ppt < 0 ] <- 0
ISD_PPT$incremental_ppt[ISD_PPT$incremental_ppt > 2 ] <- 0

plot(ISD_PPT$datetime,ISD_PPT$incremental_ppt)

# run summary stats
ts_summary(ISD_PPT$incremental_ppt)
```

## Convert hourly incremental ppt to daily
```{r}
#make hourly ppt daily - sum
ISD_PPT_daily <- ISD_PPT %>%
  mutate(date = as.Date(datetime)) %>%  # extract date
  group_by(date) %>%
  summarize(
    daily_sum = sum(incremental_ppt, na.rm = TRUE),  # sum per day
    .groups = "drop"
  )

plot(ISD_PPT_daily$date,ISD_PPT_daily$daily_sum)
```

# Visualize data

## Map CDEC stations
```{r}
#pull station locations
#ANH, ISD, BLP, DSJ, SJJ
#ANH was ANC until 2018
# ISD <- CDEC_StationInfo("ISD")
# BLP <- CDEC_StationInfo("BLP")
# SJJ <- CDEC_StationInfo("SJJ")
# CDEC_StationInfo('ANC')
# CDEC_StationInfo("DSJ")

#grab all Delta stations
library(CDECRetrieve) #had to use the following text to install: devtools::install_github("flowwest/CDECRetrieve", force = TRUE)
Delta_stations_CDEC <- cdec_stations(river_basin = "DELTA") 
Delta_stations_CDEC_sf <-  st_as_sf(Delta_stations_CDEC, coords = c("longitude", "latitude"), crs = 4326)
# Define output shapefile path
out_path <- "data/Delta_CDEC_points.shp"

# Write the sf object to a shapefile
# 'delete_dsn = TRUE' overwrites if file exists
st_write(Delta_stations_CDEC_sf, out_path, delete_dsn = TRUE)
```

```{r}
# gather shapefiles for interactive map of Delta
delta_shapefile_url <- "https://utility.arcgis.com/usrsvcs/servers/3efc635b27344a3da989ca1e7108f5e0/rest/services/Boundaries/i03_LegalDeltaBoundary/MapServer/0/query?outFields=*&where=1%3D1&f=geojson"

# Read the GeoJSON directly from the URL
# geojsonio::geojson_read returns a Spatial object
delta_legal <- st_read(delta_shapefile_url, quiet = TRUE)
# Define projection
delta_legal <- st_transform(delta_legal, crs = 4326)
```

NB some of our stations are missing...
```{r}
# interactive map of Delta with leaflet
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap tiles
  setView(lng = mean(st_coordinates(st_centroid(st_union(delta_legal)))[,1]),
          lat = mean(st_coordinates(st_centroid(st_union(delta_legal)))[,2]),
          zoom = 10) %>%
            #add legal Delta boundary
  addPolygons(
    data = delta_legal,
    color = "blue",
    weight = 2,
    opacity = 0.1,
    fillOpacity = 0.1
    #popup = ~as.character("Legal Delta") # can change to a shapefile column name
  ) %>% 
   #add the CDEC stations
  addCircleMarkers(data = Delta_stations_CDEC_sf,
  color = "red",
  weight = 1.5,
  fillOpacity = 0.4,
  popup = ~paste("Station:", station_id, name)
  )

```



# Merge data together

# Export to format needed for combining with remote sensing indices and for modeling


# Download USGS data - just in case this is useful later

- Get instantaneous USGS data (for example, discharge sensor data). Start here: ?read_waterdata_continuous. If you only need the latest value, you can use the function: ?read_waterdata_latest_continuous.

- Get daily USGS data (for example, mean daily discharge). Start here: ?read_waterdata_daily. If you only need the latest value, you can use the function: ?read_waterdata_latest_daily.

- Get discrete USGS groundwater data. Start here: ?read_waterdata_field_measurements

- Get discrete water quality data from a cooperative service that integrates publicly available water-quality data from the USGS, EPA, and over 400 state, federal, tribal, and local agencies. Start here: ?readWQPdata

- Get USGS discrete water quality data. Start here: ?read_waterdata_samples

- Get metadata about USGS time series data, including instantaneous and daily data. Start here: ?read_waterdata_ts_meta

- Find Hydro Network-Linked Data Index (NLDI) data. Start here: ?findNLDI

- Get daily data statistics. Start here: ?read_waterdata_stats_por

- Add a USGS WaterData API token to your R environment. See: https://doi-usgs.github.io/dataRetrieval/articles/read_waterdata_functions.html#api-tokens

- Get the latest status on NWIS and other data access updates. See: https://doi-usgs.github.io/dataRetrieval/articles/Status.html https://api.waterdata.usgs.gov/ogcapi/v0/collections/monitoring-locations

```{r}
# https://github.com/DOI-USGS/dataRetrieval

```

