---
title: "SURF data"
format: html
---

## Load Packages

```{r}
library(readr)
library(sf)
library(dplyr)
library(ggplot2)
library(ggspatial)
```

## Read in data

```{r}
zip_path <- tempfile(fileext = ".zip")
download.file("https://www.cdpr.ca.gov/wp-content/uploads/2025/06/SURF-Water-Data-2025.zip", zip_path, mode = "wb")

SURF_raw <- readr::read_csv(unz(zip_path, "SURF Water Data 2025.csv"))
```

## Filter to data in the Delta

```{r}
#create sf object
SURF_sf <- SURF_raw %>% 
    st_as_sf(coords = c('Longitude', 'Latitude'),
             crs = 4326,
             remove = FALSE)

#read in shapefile
delta_sf <- st_read("https://services5.arcgis.com/netd7RBEFAa05HUC/arcgis/rest/services/Legal_Delta/FeatureServer/1/query?outFields=*&where=1%3D1&f=geojson")

plot(delta_sf)
st_crs(delta_sf)

#find sampling locations within the Delta
SURF_delta_sf <- st_filter(SURF_sf, delta_sf, .predicate = st_within)
```

```{r}
##### as above but adding Suisun to legal Delta
# read in geoJSON for Delta + Suisun
delta_suisun_sf <- st_read(
  "https://services5.arcgis.com/netd7RBEFAa05HUC/ArcGIS/rest/services/LegalDelta/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson")

# checks
st_crs(SURF_sf) == st_crs(delta_suisun_sf)
plot(delta_suisun_sf)

# find sampling locations in Delta + Suisun
SURF_delta_suisun_sf <- st_filter(SURF_sf, delta_suisun_sf, .predicate = st_intersects)
```

## Explore spatial coverage

```{r}
#count samples per station
samples_per_station <- SURF_delta_sf %>%
  count(Site_code, name = "n_samples")

#plot samples per station

ggplot() +
  geom_sf(data = delta_sf, fill = "#6CA6CD", color = "white", alpha = 0.25) +
  geom_sf(
    data = samples_per_station,
    aes(size = n_samples),
    color = "#E4572E",
    alpha = 0.8
  ) +
  scale_size_continuous(name = "Samples per station") +
  labs(
    title = "Unique Station Locations within Delta",
    subtitle = "Point size reflects number of samples per station",
    x = NULL, y = NULL
  ) 

```


```{r}
##### as above but adding Suisun to legal Delta
#count samples per station
samples_per_station_ds <- SURF_delta_suisun_sf %>%
  count(Site_code, name = "n_samples")

#plot samples per station

ggplot() +
  geom_sf(data = delta_suisun_sf, fill = "#6CA6CD", color = "white", alpha = 0.25) +
  geom_sf(
    data = samples_per_station_ds,
    aes(size = n_samples),
    color = "#E4572E",
    alpha = 0.8
  ) +
  scale_size_continuous(name = "Samples per station") +
  labs(
    title = "Unique Station Locations within Delta and Suisun",
    subtitle = "Point size reflects number of samples per station",
    x = NULL, y = NULL
  ) 
```

```{r}
##adding a basemap to plot 

#transform data to web mercator to match basemap
delta_suisun_3857   <- st_transform(delta_suisun_sf, 3857)
samples_3857 <- st_transform(samples_per_station_ds, 3857)

#plot samples per station with basemap
ggplot() +
  annotation_map_tile(type = "cartolight", zoom = 10) +
  geom_sf(data = delta_suisun_3857, fill = "#6CA6CD", color = "white", alpha = 0.25) +
  geom_sf(
    data = samples_3857,
    aes(size = n_samples),
    color = "#E4572E",
    alpha = 0.8
  ) +
  scale_size_continuous(name = "Samples per station") +
  labs(
    title = "Unique Station Locations within Delta and Suisun",
    subtitle = "Point size reflects number of samples per station",
    x = NULL, y = NULL
  ) 

```

```{r}
##exploration

study_count <- SURF_delta_suisun_sf |> 
  sf::st_drop_geometry() |> 
  count(Study_description, name = "n_measurements") |> 
  arrange(desc(n_measurements))
```


Very incomplete list of stuff to do (maybe) in no particular order:
- clip to Big Break
- figure out which pesticides we are interested in
- and at what concentration
- investigate by date
- see if SFEI's cd3 dataset has anything we are missing
- figure out how to put the rosm.cache into git ignore so it is not tracked
