---
title: "SURF data"
format: html
---

## Load Packages

```{r}
library(readr)
library(sf)
library(dplyr)
library(ggplot2)
library(ggspatial)
library(leaflet)
library(htmltools)
library(stringr)
```

## Read in data

```{r}
zip_path <- tempfile(fileext = ".zip")
download.file("https://www.cdpr.ca.gov/wp-content/uploads/2025/06/SURF-Water-Data-2025.zip", zip_path, mode = "wb")

SURF_raw <- readr::read_csv(unz(zip_path, "SURF Water Data 2025.csv"))
```

## Filter to data in the Delta

```{r}
#create sf object
SURF_sf <- SURF_raw %>% 
    st_as_sf(coords = c('Longitude', 'Latitude'),
             crs = 4326,
             remove = FALSE)

#read in shapefile
delta_sf <- st_read("https://services5.arcgis.com/netd7RBEFAa05HUC/arcgis/rest/services/Legal_Delta/FeatureServer/1/query?outFields=*&where=1%3D1&f=geojson")

plot(delta_sf)
st_crs(delta_sf)

#find sampling locations within the Delta
SURF_delta_sf <- st_filter(SURF_sf, delta_sf, .predicate = st_within)
```

```{r}
##### as above but adding Suisun to legal Delta
# read in geoJSON for Delta + Suisun
delta_suisun_sf <- st_read(
  "https://services5.arcgis.com/netd7RBEFAa05HUC/ArcGIS/rest/services/LegalDelta/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson")

# checks
st_crs(SURF_sf) == st_crs(delta_suisun_sf)
plot(delta_suisun_sf)

# find sampling locations in Delta + Suisun
SURF_delta_suisun_sf <- st_filter(SURF_sf, delta_suisun_sf, .predicate = st_intersects)
```

## Explore spatial coverage

```{r}
#count samples per station
samples_per_station <- SURF_delta_sf %>%
  count(Site_code, name = "n_samples")

#plot samples per station

ggplot() +
  geom_sf(data = delta_sf, fill = "#6CA6CD", color = "white", alpha = 0.25) +
  geom_sf(
    data = samples_per_station,
    aes(size = n_samples),
    color = "#E4572E",
    alpha = 0.8
  ) +
  scale_size_continuous(name = "Samples per station") +
  labs(
    title = "Unique Station Locations within Delta",
    subtitle = "Point size reflects number of samples per station",
    x = NULL, y = NULL
  ) 

```


```{r}
##### as above but adding Suisun to legal Delta
#count samples per station
samples_per_station_ds <- SURF_delta_suisun_sf %>%
  count(Site_code, name = "n_samples")

#plot samples per station

ggplot() +
  geom_sf(data = delta_suisun_sf, fill = "#6CA6CD", color = "white", alpha = 0.25) +
  geom_sf(
    data = samples_per_station_ds,
    aes(size = n_samples),
    color = "#E4572E",
    alpha = 0.8
  ) +
  scale_size_continuous(name = "Samples per station") +
  labs(
    title = "Unique Station Locations within Delta and Suisun",
    subtitle = "Point size reflects number of samples per station",
    x = NULL, y = NULL
  ) 
```

```{r}
##adding a basemap to plot 

#transform data to web mercator to match basemap
delta_suisun_3857   <- st_transform(delta_suisun_sf, 3857)
samples_3857 <- st_transform(samples_per_station_ds, 3857)

#plot samples per station with basemap
ggplot() +
  annotation_map_tile(type = "cartolight", zoom = 10) +
  geom_sf(data = delta_suisun_3857, fill = "#6CA6CD", color = "white", alpha = 0.25) +
  geom_sf(
    data = samples_3857,
    aes(size = n_samples),
    color = "#E4572E",
    alpha = 0.8
  ) +
  scale_size_continuous(name = "Samples per station") +
  labs(
    title = "Unique Station Locations within Delta and Suisun",
    subtitle = "Point size reflects number of samples per station",
    x = NULL, y = NULL
  ) 

```

```{r}
##exploration

study_count <- SURF_delta_suisun_sf |> 
  sf::st_drop_geometry() |> 
  count(Study_description, name = "n_measurements") |> 
  arrange(desc(n_measurements))
```


```{r}
## Looking at pesticide types

#this reduces pesticide count from 505 to 178
SURF_ds_detects <- SURF_delta_suisun_sf %>%
  filter(Concentration_ppb > 0)

# export as CSV for analysis
chemicals_detected <- unique(SURF_ds_detects$Chemical_name)
write.csv(
  data.frame(chemical_name = chemicals_detected),
  "./data/raw/SURF_ds_chemicals_detected.csv",
  row.names = FALSE
)

# # export csv for analysis [this includes pesticides with ppb 0.0]
# chemical_names <- unique(SURF_delta_suisun_sf$Chemical_name)
# write.csv(
#   data.frame(chemical_name = chemical_names),
#   "./data/raw/SURF_ds_chemicals.csv",
#   row.names = FALSE
# )
```

## Shaela's workspace
```{r}
#number of measurements per chemical and per location

#dataframe to hold coords
locations_df <- SURF_ds_detects %>%
  distinct(Site_code, Latitude, Longitude)

#count chemical per site, filter, and attach coords
chem_loc_counts <- SURF_ds_detects %>%
  count(Chemical_name, Site_code, name = "num_measurements") %>%
  filter(num_measurements >= 1) |> 
  left_join(locations_df, by = "Site_code")

chem_loc_counts

#list of unique chemicals
chem_list <- chem_loc_counts |> 
  distinct(Chemical_name)

#55 total chemicals with more than 20 measurements at one location

site_07177 <- SURF_raw |> 
  filter(Site_code == "07_177")
```

```{r}
#plotting chem/loc data from above

#restructuring df to one row per Site
site_summary <- chem_loc_counts %>%
  group_by(Site_code) %>%
  summarize(
    Latitude = first(Latitude),
    Longitude = first(Longitude),
    total_measurements = sum(num_measurements),
    chem_list = str_c(
      Chemical_name, ": ", num_measurements,
      collapse = "<br>"
    ),
    .groups = "drop"
  )

#mapping
leaflet(site_summary) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
    lng = ~Longitude, lat = ~Latitude,
    radius = ~pmin(4 + sqrt(total_measurements), 18),
    fillColor = "#3182bd", fillOpacity = 0.7,
    stroke = TRUE, color = "white", weight = 1,
    popup = ~paste0(
      "<b>Site: </b>", Site_code, "<br>",
      "<b>Total measurements: </b>", total_measurements, "<br><br>",
      "<b>Chemicals:</b><br>", chem_list
    )
  )



```




















## Nick's workspace
```{r}

```

Very incomplete list of stuff to do (maybe) in no particular order:
- clip to Big Break
- investigate by date
- add number of studies to the chemical count
- add time period for each chemical - start date and end date
- figure out which pesticides we are interested in
- and at what concentration
- see if SFEI's cd3 dataset has anything we are missing
- figure out how to put the rosm.cache into git ignore so it is not tracked
- 

