---
title: "SURF data"
format: html
---

# Load Packages

```{r}
library(readr)
library(sf)
library(dplyr)
library(ggplot2)
library(ggspatial)
library(leaflet)
library(htmltools)
library(stringr)
library(readxl)
library(janitor)
library(prettymapr)
```

# Read in and wrangle data

## deprecated data
This data has apparently been deprecated and superseded by 3 different dated datasets (see below)
#```{r}
zip_path <- tempfile(fileext = ".zip")
download.file("https://www.cdpr.ca.gov/wp-content/uploads/2025/06/SURF-Water-Data-2025.zip", zip_path, mode = "wb")

SURF_old_raw <- readr::read_csv(unz(zip_path, "SURF Water Data 2025.csv"))
#```

## 1950-2009 SURF Data
```{r}
# Set paths
SURF_2009_url <- "https://www.cdpr.ca.gov/wp-content/uploads/2026/06/surf_water_data_1950-2009.xlsx"
data_dir <- "data"
SURF_2009_file <- file.path(data_dir, "SURF_2009.xlsx")

# Download data if not already cached
if (!file.exists(SURF_2009_file)) {
  message("Downloading file...")
  dir.create(data_dir, showWarnings = FALSE, recursive = TRUE)
  download.file(
    url = SURF_2009_url,
    destfile = SURF_2009_file, 
    mode = "wb"
  )
  message("Saved to: ", SURF_2009_file)
} else {
  message("Using cached file.")
}

# Load the (locally cached) data
SURF_2009 <- read_excel(SURF_2009_file, skip = 1)
```

## 2010-2017 SURF Data
```{r}
# Set paths
SURF_2017_url <- "https://www.cdpr.ca.gov/wp-content/uploads/2026/06/surf_water_data_2010-2017.xlsx"
data_dir <- "data"
SURF_2017_file <- file.path(data_dir, "SURF_2017.xlsx")

# Download data if not already cached
if (!file.exists(SURF_2017_file)) {
  message("Downloading file...")
  dir.create(data_dir, showWarnings = FALSE, recursive = TRUE)
  download.file(
    url = SURF_2017_url,
    destfile = SURF_2017_file, 
    mode = "wb"
  )
  message("Saved to: ", SURF_2017_file)
} else {
  message("Using cached file.")
}

# Load the (locally cached) data
SURF_2017 <- read_excel(SURF_2017_file, skip = 1)
```

## 2018-2025 SURF data
```{r}
# Set paths
SURF_2025_url <- "https://www.cdpr.ca.gov/wp-content/uploads/2026/06/surf_water_data_2018-2025.xlsx"
data_dir <- "data"
SURF_2025_file <- file.path(data_dir, "SURF_2025.xlsx")

# Download data if not already cached
if (!file.exists(SURF_2025_file)) {
  message("Downloading file...")
  dir.create(data_dir, showWarnings = FALSE, recursive = TRUE)
  download.file(
    url = SURF_2025_url,
    destfile = SURF_2025_file, 
    mode = "wb"
  )
  message("Saved to: ", SURF_2025_file)
} else {
  message("Using cached file.")
}

# Load the (locally cached) data
SURF_2025 <- read_excel(SURF_2025_file, skip = 1)
```

## Bind three dataframes (2009, 2017, 2025) together

```{r}
#fix column data type issue
SURF_2009 <- SURF_2009 %>% mutate(`Method Detection Level PPB` = readr::parse_double(as.character(`Method Detection Level PPB`)))
SURF_2017 <- SURF_2017 %>% mutate(`Method Detection Level PPB` = readr::parse_double(as.character(`Method Detection Level PPB`)))
SURF_2025 <- SURF_2025 %>% mutate(`Method Detection Level PPB` = readr::parse_double(as.character(`Method Detection Level PPB`)))

#combine all data
SURF_all <- bind_rows(SURF_2009, SURF_2017, SURF_2025)

#fix column names to snake case
SURF_all <- SURF_all |> clean_names()

#convert NAs in "chemical_name" column to actual missing values
SURF_all <- SURF_all %>%
  mutate(
    chemical_name = chemical_name |> 
      str_trim() |>              # remove leading/trailing spaces
      na_if("NA") |>             # turn literal "NA" into real NA
      na_if("N/A") |>            # also handle common variants
      na_if("")                  # empty strings -> NA
  )

#delete any duplicate rows where all column values match (Matt indicated there may be duplicate DRMP data)
SURF_all <- SURF_all %>%
  distinct()

#convert sample_date column to date type format
SURF_all <- SURF_all |> 
  mutate(sample_date = as.Date(sample_date))
```

## !!!Filter to data in the Delta
!!!This chunk is where we need to come back to see if we can find the dropped data!!!
```{r}
#remove any rows with NAs in lat or long
SURF_all <- SURF_all %>%
  filter(!is.na(latitude) & !is.na(longitude))

#create sf object
SURF_sf <- SURF_all %>% 
    st_as_sf(coords = c('longitude', 'latitude'),
             crs = 4326,
             remove = FALSE)

# read in geoJSON for Delta + Suisun
delta_suisun_sf <- st_read(
  "https://services5.arcgis.com/netd7RBEFAa05HUC/ArcGIS/rest/services/LegalDelta/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson")

# checks
st_crs(SURF_sf) == st_crs(delta_suisun_sf)
plot(delta_suisun_sf)

# find measurements in Delta + Suisun
SURF_delta_suisun_sf <- st_filter(SURF_sf, delta_suisun_sf, .predicate = st_intersects)

#remove any rows with NA in the chemical_name BUT NEED TO COME BACK AND SEE IF WE CAN SUPPLEMENT THIS DATA
SURF_delta_suisun_sf <- SURF_delta_suisun_sf %>%
  filter(!is.na(chemical_name))
```

# General exploration

## Explore spatial coverage

```{r}
#count samples per station
samples_per_station <- SURF_delta_suisun_sf %>%
  count(site_code, name = "n_samples")

##adding a basemap to plot 

#transform data to web mercator to match basemap
delta_suisun_3857   <- st_transform(delta_suisun_sf, 3857)
samples_3857 <- st_transform(samples_per_station, 3857)

#plot samples per station with basemap
ggplot() +
  annotation_map_tile(type = "cartolight", zoom = 10) +
  geom_sf(data = delta_suisun_3857, fill = "#6CA6CD", color = "white", alpha = 0.25) +
  geom_sf(
    data = samples_3857,
    aes(size = n_samples),
    color = "#E4572E",
    alpha = 0.8
  ) +
  scale_size_continuous(name = "Samples per station") +
  labs(
    title = "Unique Station Locations within Delta and Suisun",
    subtitle = "Point size reflects number of samples per station",
    x = NULL, y = NULL
  ) 
```

## Study counts
```{r}

study_count <- SURF_delta_suisun_sf |> 
  sf::st_drop_geometry() |> 
  count(study_description, name = "n_measurements") |> 
  arrange(desc(n_measurements))
```

## Pesticide types
```{r}

#filter to data where there is a detect
SURF_ds_detects <- SURF_delta_suisun_sf %>%
  filter(concentration_ppb > 0)

# export as CSV for analysis
#chemicals_detected <- unique(SURF_ds_detects$Chemical_name)
#write.csv(
#  data.frame(chemical_name = chemicals_detected),
#  "./data/raw/SURF_ds_chemicals_detected.csv",
#  row.names = FALSE
#)
```

# Shaela's workspace

## filtering approach #1
```{r}
#number of measurements per chemical and per location

#dataframe to hold coords
locations_df <- SURF_ds_detects %>%
  distinct(site_code, latitude, longitude)

#count chemical per site, filter, and attach coords
chem_loc_counts <- SURF_ds_detects %>%
  count(chemical_name, site_code, name = "num_measurements") %>%
  filter(num_measurements >= 10) |> 
  left_join(locations_df, by = "site_code")

chem_loc_counts

#list of unique chemicals
chem_list <- chem_loc_counts |> 
  distinct(chemical_name)

#81 total chemicals with more than 10 measurements at one location
```

## filtering approach #2

```{r}
# 1) Find (chemical, site) combinations with ≥10 measurements
site_chem_ok <- SURF_ds_detects %>%
  st_drop_geometry() %>%
  count(chemical_name, site_code, name = "num_measurements") %>%
  filter(num_measurements >= 10)

# 2) Among those, keep chemicals that appear at ≥3 distinct sites
chem_ok <- site_chem_ok %>%
  count(chemical_name, name = "num_sites") %>%
  filter(num_sites >= 3)

# 3) Filter the original data to only rows that meet both criteria
SURF_filtered <- SURF_ds_detects %>%
  semi_join(site_chem_ok, by = c("chemical_name", "site_code")) %>%
  semi_join(chem_ok,     by = "chemical_name")
```

## Looking into fluridone (herbicide where lab data shows it make have impact on microcystis)

```{r}
#filter data to fluridone and sites with more than 10 measurements
fluridone_counts <- SURF_ds_detects %>%
  filter(chemical_name == "fluridone") %>%   
  add_count(site_code, name = "num_measurements") |> 
  filter(num_measurements >= 10)
```

```{r}
#fluridone data

#restructuring df to one row per Site
fl_site_summary <- fluridone_counts %>%
  group_by(site_code) %>%
  summarize(
    latitude = first(latitude),
    longitude = first(longitude),
    total_measurements = first(num_measurements),
    chem_list = first(chemical_name),
    earliest_date = min(sample_date, na.rm = TRUE),
    latest_date = max(sample_date, na.rm = TRUE),
    .groups = "drop"
  )

#mapping
leaflet(fl_site_summary) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
    lng = ~longitude, lat = ~latitude,
    radius = ~pmin(4 + sqrt(total_measurements), 18),
    fillColor = "#3182bd", fillOpacity = 0.7,
    stroke = TRUE, color = "white", weight = 1,
    popup = ~paste0(
      "<b>Site: </b>", site_code, "<br>",
      "<b>Total measurements: </b>", total_measurements, "<br><br>",
      "<b>Earliest sample date: </b>", format(earliest_date, "%Y-%m-%d"), "<br>",
      "<b>Latest sample date: </b>", format(latest_date, "%Y-%m-%d"), "<br><br>",
      "<b>Chemicals:</b><br>", chem_list
    )
  )
```

## Looking into glyphosate

```{r}
#filter data to glyphosate and sites with more than 10 measurements
glyphosate_counts <- SURF_ds_detects %>%
  filter(chemical_name == "glyphosate") %>%   
  add_count(site_code, name = "num_measurements") |> 
  filter(num_measurements >= 10)
```

```{r}
#glyphosate data

#restructuring df to one row per Site
gl_site_summary <- glyphosate_counts %>%
  group_by(site_code) %>%
  summarize(
    latitude = first(latitude),
    longitude = first(longitude),
    total_measurements = first(num_measurements),
    chem_list = first(chemical_name),
    earliest_date = min(sample_date, na.rm = TRUE),
    latest_date = max(sample_date, na.rm = TRUE),
    .groups = "drop"
  )

#mapping
leaflet(gl_site_summary) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
    lng = ~longitude, lat = ~latitude,
    radius = ~pmin(4 + sqrt(total_measurements), 18),
    fillColor = "#3182bd", fillOpacity = 0.7,
    stroke = TRUE, color = "white", weight = 1,
    popup = ~paste0(
      "<b>Site: </b>", site_code, "<br>",
      "<b>Total measurements: </b>", total_measurements, "<br><br>",
      "<b>Earliest sample date: </b>", format(earliest_date, "%Y-%m-%d"), "<br>",
      "<b>Latest sample date: </b>", format(latest_date, "%Y-%m-%d"), "<br><br>",
      "<b>Chemicals:</b><br>", chem_list
    )
  )
```

## Looking into diuron

```{r}
#filter data to diuron and sites with more than 10 measurements
diuron_counts <- SURF_ds_detects %>%
  filter(chemical_name == "diuron") %>%   
  add_count(site_code, name = "num_measurements") |> 
  filter(num_measurements >= 10) 
```

```{r}
#diuron data

#restructuring df to one row per Site
di_site_summary <- diuron_counts %>%
  group_by(site_code) %>%
  summarize(
    latitude = first(latitude),
    longitude = first(longitude),
    total_measurements = first(num_measurements),
    chem_list = first(chemical_name),
    earliest_date = min(sample_date, na.rm = TRUE),
    latest_date = max(sample_date, na.rm = TRUE),
    .groups = "drop"
  )

#mapping
leaflet(di_site_summary) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
    lng = ~longitude, lat = ~latitude,
    radius = ~pmin(4 + sqrt(total_measurements), 18),
    fillColor = "#3182bd", fillOpacity = 0.7,
    stroke = TRUE, color = "white", weight = 1,
    popup = ~paste0(
      "<b>Site: </b>", site_code, "<br>",
      "<b>Total measurements: </b>", total_measurements, "<br><br>",
      "<b>Earliest sample date: </b>", format(earliest_date, "%Y-%m-%d"), "<br>",
      "<b>Latest sample date: </b>", format(latest_date, "%Y-%m-%d"), "<br><br>",
      "<b>Chemicals:</b><br>", chem_list
    )
  )
```

```{r}
#plotting chem/loc data from above

#restructuring df to one row per Site
site_summary <- chem_loc_counts %>%
  group_by(site_code) %>%
  summarize(
    Latitude = first(latitude),
    Longitude = first(longitude),
    total_measurements = sum(num_measurements),
    chem_list = str_c(
      chemical_name, ": ", num_measurements,
      collapse = "<br>"
    ),
    .groups = "drop"
  )

#mapping
leaflet(site_summary) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
    lng = ~Longitude, lat = ~Latitude,
    radius = ~pmin(4 + sqrt(total_measurements), 18),
    fillColor = "#3182bd", fillOpacity = 0.7,
    stroke = TRUE, color = "white", weight = 1,
    popup = ~paste0(
      "<b>Site: </b>", site_code, "<br>",
      "<b>Total measurements: </b>", total_measurements, "<br><br>",
      "<b>Chemicals:</b><br>", chem_list
    )
  )
```




















# Nick's workspace
```{r}

```

Very incomplete list of stuff to do (maybe) in no particular order:
- clip to Big Break
- investigate by date
- add number of studies to the chemical count
- add time period for each chemical - start date and end date
- figure out which pesticides we are interested in
  - data from https://www.cdpr.ca.gov/pesticide-use-in-california/pesticide-use-reporting/#PUR-Data-Files-Used-to-Create-the-Annual-Reports can be used to figure out application amounts??
- and at what concentration
- see if SFEI's cd3 dataset has anything we are missing
- figure out how to put the rosm.cache into git ignore so it is not tracked
- 

