---
title: "ScienceBase Chlorophyll Data Harmonization"
format: html
Author: Emily Richardson
---

### The raw data script that pulled data from ScienceBase is within 01_chl_wrangle.qmd. Since these data were numerous and from multiple special projects, they require lengthy data harmonization. This notebook is specific to ScienceBase data manipulation so that these "massaging" techniques don't over-clutter the chlorophyll data wrangling notebook.

```{r}
#Load packages
library(janitor)
library(dplyr)
library(stringr)
library(lubridate)
```

# Three similar data sources containing toxins, qpcr, and chl data

```{r}
#Read in data
chl_gen_cyano_1 <- read.csv("data//raw//sciencebase//5c13d214e4b006c4f8532a17//Cyanotoxins_Chl_Genetic_Data.csv")
chl_gen_cyano_2 <- read.csv("data//raw//sciencebase//628cc9c2d34ef70cdba3bbff//2019_Cyanotoxins_Chl_Genetics_Data.csv", header = TRUE, skip = 1) #1st row is garbage
chl_gen_cyano_3 <- read.csv("data//raw//sciencebase//601827aad34edf5c66efffc8//2018_Cyanotoxins_Chl_Genetics_Data.csv", header = TRUE, skip = 1)#1st row is garbage

#Limit to just our Delta location and chl data (assuming toxins will be captures in VK data project)
chl_gen_cyano_1_CA <- chl_gen_cyano_1 %>% filter(State == "CA") %>%
    select(Station.ID, Station.Name, Date, Time, Sample.Type, Chlorophyll.a..ug.L.) %>%
    rename(station_num = Station.ID, station = Station.Name, date = Date, time = Time, sample_type = Sample.Type, chla_ugl = Chlorophyll.a..ug.L.)%>% filter(sample_type == "Regular")

chl_gen_cyano_1_CA$time_hms <- str_c(chl_gen_cyano_1_CA$time, ":00")
chl_gen_cyano_1_CA <- chl_gen_cyano_1_CA %>% select(-time)
chl_gen_cyano_1_CA$datetime <- paste(chl_gen_cyano_1_CA$date, chl_gen_cyano_1_CA$time, sep = " ")


chl_gen_cyano_2_CA <- chl_gen_cyano_2 %>% filter(str_detect(USGS.Station.Name, "CA"))%>%
    select(USGS.Station.ID, USGS.Station.Name, Sample.Date, Sample.Time, Sample.Type, Chlorophyll.a..ug.L.)%>%
    rename(station_num = USGS.Station.ID, station = USGS.Station.Name, date = Sample.Date, time = Sample.Time, sample_type = Sample.Type, chla_ugl = Chlorophyll.a..ug.L.)%>% filter(sample_type == "Regular")

chl_gen_cyano_2_CA$time_hm <- sub("(?=..$)", ":", chl_gen_cyano_2_CA$time, perl = TRUE)
chl_gen_cyano_2_CA$time_hms <- str_c(chl_gen_cyano_2_CA$time_hm, ":00")
chl_gen_cyano_2_CA <- chl_gen_cyano_2_CA %>% select(-time, -time_hm)
chl_gen_cyano_2_CA$datetime <- paste(chl_gen_cyano_2_CA$date, chl_gen_cyano_2_CA$time, sep = " ")

chl_gen_cyano_3_CA <- chl_gen_cyano_3 %>% filter(str_detect(USGS.Station.Name, "CA")) %>%
    select(USGS.Station.ID, USGS.Station.Name, Sample.Date, Sample.Time, Sample.Type, Chlorophyll.a..ug.L.)%>%
    rename(station_num = USGS.Station.ID, station = USGS.Station.Name, date = Sample.Date, time = Sample.Time, sample_type = Sample.Type, chla_ugl = Chlorophyll.a..ug.L.)%>% filter(sample_type == "Regular")

chl_gen_cyano_3_CA$time_hm <- sub("(?=..$)", ":", chl_gen_cyano_3_CA$time, perl = TRUE)
chl_gen_cyano_3_CA$time_hms <- str_c(chl_gen_cyano_3_CA$time_hm, ":00")
chl_gen_cyano_3_CA <- chl_gen_cyano_3_CA %>% select(-time, -time_hm)
chl_gen_cyano_3_CA$datetime <- paste(chl_gen_cyano_3_CA$date, chl_gen_cyano_3_CA$time, sep = " ")

freeport_chl_ds <-rbind(chl_gen_cyano_1_CA, chl_gen_cyano_2_CA, chl_gen_cyano_3_CA)
rm(chl_gen_cyano_1, chl_gen_cyano_2, chl_gen_cyano_3, chl_gen_cyano_1_CA, chl_gen_cyano_2_CA, chl_gen_cyano_3_CA)

freeport_chl_ds$agency <- "USGS"
freeport_chl_ds$latitude <- 38.45566
freeport_chl_ds$longitude <- -121.50162


freeport_chl_ds <- freeport_chl_ds %>% mutate(datetime = mdy_hms(datetime))%>%
    select(-time_hms, -sample_type) %>%
    mutate(date = mdy(date))

freeport_chl_ds$chla_ugl <- as.double(freeport_chl_ds$chla_ugl)
freeport_chl_ds <- freeport_chl_ds %>% filter(!is.na(chla_ugl))
freeport_chl_ds$sample_type <- "discrete"
```

# USGS high resolution - Many datasets using the same boat setup and instrumentation can be combined
```{r}
hr_mapping2018_jul <- read.csv("data//raw//sciencebase//5d3f1ae2e4b01d82ce8d90e0//Delta Water Quality Mapping July 2018 High Resolution.csv")
hr_mapping2018_may <- read.csv("data//raw//sciencebase//5d3f1ae2e4b01d82ce8d90e0//Delta Water Quality Mapping May 2018 High Resolution.csv")
hr_mapping2018_oct <-read.csv("data//raw//sciencebase//5d3f1ae2e4b01d82ce8d90e0//Delta Water Quality Mapping October 2018 High Resolution.csv")
ds_mapping2018 <-read.csv("data//raw//sciencebase//5d3f1ae2e4b01d82ce8d90e0//Delta Water Quality Mapping 2018 Discrete Sampling v2.0.csv")

hr_toe20180823 <- read.csv("data//raw//sciencebase//5e7ceb88e4b01d5092751c18//2018-08-23_Sample_data_ver.2.0.csv")
hr_toe20180905 <- read.csv("data//raw//sciencebase//5e7ceb88e4b01d5092751c18//2018-09-05_Sample_data_ver.2.0.csv")
hr_toe20180913 <- read.csv("data//raw//sciencebase//5e7ceb88e4b01d5092751c18//2018-09-13_Sample_data_ver.2.0.csv")
hr_toe20181004 <- read.csv("data//raw//sciencebase//5e7ceb88e4b01d5092751c18//2018-10-04_Sample_data_ver.2.0.csv")

hrprof_aqveg2018 <- read.csv("data//raw//sciencebase//5f1f44c582cef313ed8ec753//AquaticVegetation_2018.csv")

hrds_conf2017 <- read.csv("data//raw//sciencebase//5f5fe6eb82ce3550e3bff4a5//Confluence_2017.csv")

hr_yolo2017 <- read.csv("data//raw//sciencebase//5f5ff87582ce3550e3bff547//YoloBypass_2017.csv")

#ds_opbwrt <- read.csv("data//raw//sciencebase//5f6ce12a82ce38aaa24801be//OpB WRT Discrete.csv")
#hr_opbwrt_2017oct <- read.csv("data//raw//sciencebase//5f6ce12a82ce38aaa24801be//OpB WRT Oct2017_HR.csv")
#hr_opbwrt_2018may <- read.csv("data//raw//sciencebase//5f6ce12a82ce38aaa24801be//OpB WRT May2018 HR.csv")
#hr_opbwrt_2018oct <- read.csv("data//raw//sciencebase//5f6ce12a82ce38aaa24801be//OpB WRT Oct2018_HR.csv")

hr_edb <- read.csv("data//raw//sciencebase//5f7dee6482ce1d74e7dd9da0//EDB_HR.txt")
ds_edb <- read.csv("data//raw//sciencebase//5f7dee6482ce1d74e7dd9da0//EDB_DS.txt")


hr_csc_2021Jul <- read.csv("data//raw//sciencebase//5f89cbea82ce3241878937f6//July2021_CacheSloughComplexMapping_HR_preliminary.txt")
hr_csc_2021Sept <- read.csv("data//raw//sciencebase//5f89cbea82ce3241878937f6//September2021_CacheSloughComplexMapping_HR_preliminary.txt")
hr_csc_2022Oct <- read.csv("data//raw//sciencebase//5f89cbea82ce3241878937f6//October2022_CacheSloughComplexMapping_HR_preliminary.txt")
ds_csc <- read.csv("data//raw//sciencebase//5f89cbea82ce3241878937f6//CacheSloughComplexMapping_DS_preliminary.txt")

hr_ftswq <- read.csv("data//raw//sciencebase//5f452e8782ce4c3d12251702//FTS_water_quality_20s_median.csv")

ds_mapping2020 <-read.csv("data//raw//sciencebase/5f24747f82cef313ed971974//DeltaMapping_2020_DS.txt")
ds_mapping2021 <-read.csv("data//raw//sciencebase//5f24747f82cef313ed971974//DeltaMapping_2021_DS.txt")
hr_mapping2020aug <-read.csv("data//raw//sciencebase//5f24747f82cef313ed971974//DeltaMapping_August2020_HR.txt")
hr_mapping2021jul <-read.csv("data//raw//sciencebase//5f24747f82cef313ed971974//DeltaMapping_July2021_HR.txt")
hr_mapping2020jul <-read.csv("data//raw//sciencebase//5f24747f82cef313ed971974//DeltaMapping_July2020_HR.txt")
hr_mapping2021may <-read.csv("data//raw//sciencebase//5f24747f82cef313ed971974//DeltaMapping_May2021_HR.txt")
hr_mapping2020june <-read.csv("data//raw//sciencebase//5f24747f82cef313ed971974//DeltaMapping_June2020_HR.txt")
hr_mapping2021oct <-read.csv("data//raw//sciencebase//5f24747f82cef313ed971974//DeltaMapping_October2021_HR.txt")
hr_mapping2020oct <-read.csv("data//raw//sciencebase//5f24747f82cef313ed971974//DeltaMapping_October2020_HR.txt")

ds_srinc2019 <- read.csv("data//raw//sciencebase//5fa096b1d34e198cb7919341//SRiNC_2019_DS.txt")
hr_srinc2019 <- read.csv("data//raw//sciencebase//5fa096b1d34e198cb7919341//SRiNC_2019_HR.txt")

ds_frank <-read.csv("data//raw//sciencebase//64ee7f82d34e4b6d8ab71438//FranksTract_2022_DS.txt")
hr_frank <-read.csv("data//raw//sciencebase//64ee7f82d34e4b6d8ab71438//FranksTract_2022-2023_HR.txt")

ds_mapping2022 <-read.csv("data//raw//sciencebase//64fa6804d34ed30c205504a6//DeltaMapping_2022_DS.txt")
hr_mapping2022jul <-read.csv("data//raw//sciencebase//64fa6804d34ed30c205504a6//DeltaMapping_July2022_HR.txt")
hr_mapping2022may <-read.csv("data//raw//sciencebase//64fa6804d34ed30c205504a6//DeltaMapping_May2022_HR.txt")
hr_mapping2022oct <-read.csv("data//raw//sciencebase//64fa6804d34ed30c205504a6//DeltaMapping_October2022_HR.txt")

ds_sensor_hyperspec <- read.csv("data//raw//sciencebase//65b18ae7d34e36a39044c493//T6_DS_DR.csv")

ds_mapping2023 <- read.csv("data//raw//sciencebase//6596fe37d34e3265ab158d60//DeltaMapping_2023_DS.txt")
hr_mapping2023jul <- read.csv("data//raw//sciencebase//6596fe37d34e3265ab158d60//DeltaMapping_July2023_HR.txt")
hr_mapping2023may <- read.csv("data//raw//sciencebase//6596fe37d34e3265ab158d60//DeltaMapping_May2023_HR.txt")
hr_mapping2023oct  <- read.csv("data//raw//sciencebase//6596fe37d34e3265ab158d60//DeltaMapping_October2023_HR.txt")

hr_mapping2024july <- read.csv("data//raw//sciencebase//66427d79d34e1955f5a41cf0//HR_DeltaMapping_July2024_provisional.txt")
hr_mapping2024may <- read.csv("data//raw//sciencebase//66427d79d34e1955f5a41cf0//HR_DeltaMapping_May2024_provisional.txt")
hr_mapping2024oct <- read.csv("data//raw//sciencebase//66427d79d34e1955f5a41cf0//HR_DeltaMapping_October2024_provisional.txt")
```

## SF Bay Peterson data (has a Delta extent -Grizzly Bay to Rio Vista in Sac River)

```{r}
ds_peter20162021 <- read.csv("data//raw//sciencebase//5966abe6e4b0d1f9f05cf551//2016_2021SanFranciscoBayWaterQualityData.csv")
peterlocs20162021 <- read.csv("data//raw//sciencebase//5966abe6e4b0d1f9f05cf551//SFBay_TableofStationLocations_v4.csv")

ds_peter19692015 <- read.csv("data//raw//sciencebase//64248ee5d34e370832fe343d//SanFranciscoBayWaterQualityData1969-2015v4.csv")
peterlocs19692015 <- read.csv("data//raw//sciencebase//64248ee5d34e370832fe343d//SFBstation_locations19692015.csv")
```

## Graham Phyto stdy

```{r}
graham_phyto <- read.csv("data//raw//sciencebase//601025fcd34e162231fecfca//2018 Phytoplankton Data.csv")
```
