---
title: "Chlorophyll Data integration and explotraion"
format: html
Author: Emily Richardson, TBD
---
#DO NOT RUN IT WILL EXPLODE YOUR COMPUTER. I messed up the map rendering and have far too many points to disply., I need to aggregate further.
# This workbook contains code to integrate pre-cleaned chlorophyll data. See 01-prefixed notebooks for cleaning steps.

## Load Packages
```{r}
library(dplyr)
library(lubridate)
library(ggplot2)
library(plotly)
library(sf)
library(leaflet)
library(leaflet.extras2)
library(scales)
library(crosstalk)
library(geojsonsf)
library(jsonlite)
```

## Read in data and keep only dates that fall within the range of continous monitoring data

```{r}
#Continuous monitoring
cm_data <- read.csv("data//01_processed//01_chl_cm_daytime_avg.csv")
cm_data <- cm_data %>% select(-X)
maxdate <- as.Date(max(cm_data$date))
mindate <- as.Date(min(cm_data$date))

#ScienceBase
sb_data <- read.csv("data//01_processed//01_chl_sciencebase.csv")
sb_data <- sb_data %>% 
    mutate(date = as.Date(datetime)) %>%
    filter(between(date, mindate, maxdate))

#EMP
## TBD
#EDI
## TBD

rm(mindate, maxdate)
```

## Combine the data

```{r}
cm_data <- cm_data %>% 
    select(-station, -group) %>%
    rename(sample_type = instrument,
    chl_ugl = avg_fchl) %>%
    mutate(date = as.Date(date)) %>% 
    mutate(chl_ugl = round(chl_ugl, 2)) %>% 
    distinct()

sb_data <- sb_data %>%
    select(-datetime) %>%
    distinct()

chl_data <- rbind(cm_data, sb_data)
rm(cm_data, sb_data)

chl_data <- chl_data %>% filter(!is.na(latitude)) %>% filter(!is.na(longitude)) %>% filter(!is.na(chl_ugl))
```

## Visulaize the data

### Convert and clean

```{r}
#Create spatial object
chl_sf <- chl_data %>% 
    st_as_sf(coords = c('longitude', 'latitude'),
             crs = 4326, #WGS84, leaflet demands it.
             remove = FALSE)

#read in shapefile
# read in geoJSON for Delta + Suisun
delta_suisun_sf <- st_read(
  "https://services5.arcgis.com/netd7RBEFAa05HUC/ArcGIS/rest/services/LegalDelta/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson")

nad83_delta_suisun_sf <- st_transform(delta_suisun_sf, crs = 4326)
rm(delta_suisun_sf)

# find sampling locations in Delta + Suisun
chl_delta_suisun_sf <- st_filter(chl_sf, nad83_delta_suisun_sf, .predicate = st_intersects)
rm(chl_sf)

#group by hydrologic cycle
chl_seasonal <- chl_delta_suisun_sf %>%
  mutate(
    year = year(date),
    month = month(date),
    season = case_when(
      month %in% c(10, 11, 12, 1, 2, 3) ~ "Wet",
      month %in% c(4, 5, 6, 7, 8, 9)     ~ "Dry"),
    water_year = if_else(month >= 10, year + 1, year),
    season_date = case_when(
      season == "Wet" ~ as.Date(paste0(water_year, "-01-15")),
      season == "Dry" ~ as.Date(paste0(water_year, "-07-15"))
  ))
```

### Create fodder for map

```{r}
#create labels
chl_seasonal$time_label <- paste(
  "WY", chl_seasonal$water_year,
  chl_seasonal$season)

shared_data <- SharedData$new(
  chl_seasonal,
  key = ~time_label)

filter_slider(
  id = "season",
  label = "Season",
  sharedData = shared_data,
  column = ~water_year)

leaflet(shared_data) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
    radius = ~scales::rescale(chl_ugl, c(4,12)),
    color = ~viridisLite::viridis(100)[
      cut(chl_ugl, 100)])

#create color and marker scales
pal <- colorNumeric(
  palette = "viridis",
  domain = chl_seasonal$chl_ugl)

chl_seasonal$radius <- rescale(
  chl_seasonal$chl_ugl,
  to = c(4, 12)
  )
#convert to timeline-compatible GeoJSON
geo <- fromJSON(
  sf_geojson(chl_seasonal),
  simplifyVector = FALSE)

#add timeline information and styling
for(i in seq_along(geo$features)) {

  geo$features[[i]]$properties$time <-
    format(chl_seasonal$season_date[i], "%Y-%m-%d")

  geo$features[[i]]$properties$style <- list(
    radius = chl_seasonal$radius[i],
    color = pal(chl_seasonal$chl_ugl[i]),
    fillColor = pal(chl_seasonal$chl_ugl[i]),
    fillOpacity = 0.8,
    weight = 1)

geo$features[[i]]$properties$popup <- paste0(
  "<b>", chl_seasonal$time_label[i], "</b>",
  "<br><b>Chl:</b> ",
  round(chl_seasonal$chl_ugl[i], 1),
  " µg/L")}

geojson <- toJSON(
  geo,
  auto_unbox = TRUE)
```

### Build map

```{r}
leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%

  addPolygons(
    data = nad83_delta_suisun_sf,
    color = "black",
    weight = 1,
    fill = FALSE
  ) %>%

  addTimeline(
    geojson,
    timelineOpts = timelineOptions(
      formatOutput = TRUE,
      showTicks = TRUE,
      duration = 0
    )
  ) %>%

  addLegend(
    pal = pal,
    values = chl_seasonal$chl_ugl,
    title = "Chlorophyll (µg/L)"
  )
```
