---
title: "Download and wrangle hydro datasets"
format: 
  html:
    toc: true
    number-sections: true
    embed-resources: true
authors: Michelle Stern and Nick Rowlands
---

# Brainstorm

The primary environmental factors shaping phytoplankton abundance and community composition are salinity, hydrodynamics, stratification, light, temperature, nutrients, and time. (https://essopenarchive.org/doi/pdf/10.22541/essoar.176495220.03206867/v1) \* Positive: high nutrients, low turbulence, long residence time, adequate light, warm temperatures, low grazing rates, seed population/refugia, low biodiversity. \* Modulating Factors: strong environmental gradients like stratification and salinity, nutrient frms and ratios, diverse habitats/connectivity, selective grazing, predator/prey co-evolution, adequate light, toxin production \* Negative: low nutrients\*, high turbulence, short residence time, low light, high grazing rates, high biodiversity, extreme temperatures warm or cold

Variables from SCHISM: water temperature, flow, turbidity, salinity, residence time (average for a certain month? season?) how to get an index for stratification, light? Variables from climate datasets: wind, solar radiation (also when?) Flow at certain gages? satellite data? Non time-variant variables?

# Download data

```{r}
# Required packages to include
# install.packages("librarian")
librarian::shelf(readr, dplyr, tidyr, here, leaflet, sf, ggplot2, ggspatial, prettymapr)
```

![[SCHISM boundary conditions](https://cadwrdeltamodeling.github.io/BayDeltaSCHISM/topics/baydelta_boundaries.html)](https://cadwrdeltamodeling.github.io/BayDeltaSCHISM/_images/open_boundaries.png){width="80%"}

```{r}
# SCHISM data

```

# Download other data

- Get instantaneous USGS data (for example, discharge sensor data). Start here: ?read_waterdata_continuous. If you only need the latest value, you can use the function: ?read_waterdata_latest_continuous.

- Get daily USGS data (for example, mean daily discharge). Start here: ?read_waterdata_daily. If you only need the latest value, you can use the function: ?read_waterdata_latest_daily.

- Get discrete USGS groundwater data. Start here: ?read_waterdata_field_measurements

- Get discrete water quality data from a cooperative service that integrates publicly available water-quality data from the USGS, EPA, and over 400 state, federal, tribal, and local agencies. Start here: ?readWQPdata

- Get USGS discrete water quality data. Start here: ?read_waterdata_samples

- Get metadata about USGS time series data, including instantaneous and daily data. Start here: ?read_waterdata_ts_meta

- Find Hydro Network-Linked Data Index (NLDI) data. Start here: ?findNLDI

- Get daily data statistics. Start here: ?read_waterdata_stats_por

- Add a USGS WaterData API token to your R environment. See: https://doi-usgs.github.io/dataRetrieval/articles/read_waterdata_functions.html#api-tokens

- Get the latest status on NWIS and other data access updates. See: https://doi-usgs.github.io/dataRetrieval/articles/Status.html https://api.waterdata.usgs.gov/ogcapi/v0/collections/monitoring-locations

```{r}
# https://github.com/DOI-USGS/dataRetrieval

```

# Visualize data

```{r}
# gather shapefiles for interactive map of Delta
delta_shapefile_url <- "https://utility.arcgis.com/usrsvcs/servers/3efc635b27344a3da989ca1e7108f5e0/rest/services/Boundaries/i03_LegalDeltaBoundary/MapServer/0/query?outFields=*&where=1%3D1&f=geojson"

# Read the GeoJSON directly from the URL
# geojsonio::geojson_read returns a Spatial object
delta_legal <- st_read(delta_shapefile_url, quiet = TRUE)
# Define projection
delta_legal <- st_transform(delta_legal, crs = 4326)
```

```{r}
# interactive map of Delta with leaflet
leaflet(delta_legal) %>%
  addTiles() %>%  # Add default OpenStreetMap tiles
  setView(lng = mean(st_coordinates(st_centroid(st_union(delta_legal)))[,1]),
          lat = mean(st_coordinates(st_centroid(st_union(delta_legal)))[,2]),
          zoom = 10) %>%
  addPolygons(
    color = "blue",
    weight = 1,
    opacity = 0.6,
    fillOpacity = 0.2,
    popup = ~as.character("Legal Delta") # can change to a shapefile column name
  )
```

```{r}
## syntax for adding multiple shapefiles example code
# Paths to shapefiles
shp_files <- c("data/roads.shp", "data/lakes.shp", "data/parks.shp")

# Read shapefiles into a list
shapes <- lapply(shp_files, read_sf)

# Create leaflet map
leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = shapes[[1]], color = "red", weight = 2, fillOpacity = 0.3, group = "Roads") %>%
  addPolygons(data = shapes[[2]], color = "blue", weight = 2, fillOpacity = 0.3, group = "Lakes") %>%
  addPolygons(data = shapes[[3]], color = "green", weight = 2, fillOpacity = 0.3, group = "Parks") %>%
  addLayersControl(
    overlayGroups = c("Roads", "Lakes", "Parks"),
    options = layersControlOptions(collapsed = FALSE)
  )
```

# Merge data together

# Export to format needed for modeling
